\chapter{Mention Count Source List}
\label{appendix:mention-count-source-list}

This appendix lists the sources used to construct the consolidated mention-count ranking described in Section~\ref{sec:mention-count-ranking}. Sources fall into three groups: community-curated robotics software lists, academic survey or comparison papers, and auxiliary LLM-assisted discovery lists. The LLM-assisted lists were used only as supplementary discovery aids, as discussed in Section~\ref{subsec:llm-assisted-sources}.

\section*{Community-Curated Sources}

Table~\ref{tab:community-mention-sources} lists the 15 community-curated sources used for candidate discovery. Each source is a public list, GitHub Topic page, or Wikipedia article that aggregates robotics software relevant to motion planning or simulation.
The four GitHub Topic pages are retained because they are recorded in the source catalogue in the supplementary project materials. They are dynamic GitHub aggregations, so copied page text was not preserved in the raw-material archive in the same way as the static curated-list pages.

\begin{table}[htbp]
\centering
\caption{Community-curated sources used for mention counting.}
\label{tab:community-mention-sources}
\small
\singlespacing
\begin{tabular}{@{}p{4.6cm}p{8.2cm}@{}}
\toprule
\textbf{Source} & \textbf{Description} \\
\midrule
GitHub Topic: motion-planning & GitHub Topic page aggregating active motion planning projects. \\
GitHub Topic: robotics-simulation & GitHub Topic page aggregating robotics simulation projects. \\
GitHub Topic: path-planning & GitHub Topic page for path and motion planning. \\
GitHub Topic: robotics-navigation & GitHub Topic page for navigation and planning. \\
best-of-robot-simulators & Curated list of robotic simulators (knmcguire). \\
Awesome Robotics Libraries & Robotics library list with planning and simulation sections (jslee02). \\
Awesome Robotics (ahundt) & Comprehensive robotics list covering planning, optimization, and simulation. \\
Awesome Robotics (kiloreux) & Community-maintained robotics list with planning and simulation entries. \\
Awesome Motion Planning & Motion planning list covering classical and modern libraries (AGV-IIT-KGP). \\
Awesome Multibody Dynamics Simulation & Multibody dynamics and simulation list (jslee02). \\
Awesome Robotic Tooling & Robotic tools and libraries list (Ly0n). \\
Awesome ROS 2 & ROS~2 ecosystem list including navigation, planning, and simulation (fkromer). \\
Awesome Webots & Webots-focused ecosystem list (lukicdarkoo). \\
Awesome Robotics Projects & Robotics projects list with planning and simulation components (mjyc). \\
Wikipedia: Robotics simulator & Wikipedia article aggregating robotics simulator entries. \\
\bottomrule
\end{tabular}
\end{table}

\section*{Academic Sources}

Table~\ref{tab:academic-mention-sources} lists the ten academic survey and comparison papers used for candidate discovery, with full bibliographic entries in the reference list. These papers survey or compare simulators, physics engines, and development frameworks used in robotics research; the second column gives representative packages from the selected set that each paper discusses.

\begin{table}[htbp]
\centering
\caption{Academic survey and comparison papers used for mention counting.}
\label{tab:academic-mention-sources}
\small
\singlespacing
\begin{tabular}{@{}p{8.6cm}p{4.2cm}@{}}
\toprule
\textbf{Academic source} & \textbf{Example packages discussed} \\
\midrule
\citet{Brugali2007}, ``Trends in Robot Software Domain Engineering'' (book chapter) & ROS, OROCOS, OpenRTM-aist, Gazebo, YARP \\
\citet{Zlajpah2008}, ``Simulation in Robotics'' & GraspIt!, ODE, Newton Dynamics, CoppeliaSim \\
\citet{Reckhaus2010}, ``An Overview about Simulation and Emulation in Robotics'' & GraspIt!, OpenRAVE, Gazebo \\
\citet{Harris2011}, ``Survey of Popular Robotics Simulators, Frameworks, and Toolkits'' & Gazebo, ROS, OpenRAVE, MRPT, Webots \\
\citet{Staranowicz2011}, ``A Survey and Comparison of Commercial and Open-Source Robotic Simulator Software'' & Webots, Gazebo \\
\citet{InigoBlasco2012}, ``Robotics Software Frameworks for Multi-Agent Robotic Systems Development'' & ROS, OROCOS, OpenRTM-aist, Gazebo, YARP \\
\citet{collins2021simulators}, ``A Review of Physics Simulators for Robotic Applications'' & Gazebo, CoppeliaSim, Webots, MuJoCo, CARLA, SOFA, Project Chrono \\
\citet{Korber2021}, ``Comparing Popular Simulation Environments in the Scope of Robotics and Reinforcement Learning'' & Gazebo, MuJoCo, PyBullet, Webots, DART, Simbody, PhysX \\
\citet{Farley2022}, ``How to Pick a Mobile Robot Simulator'' & Gazebo, Webots, MORSE, CoppeliaSim, ODE, DART, Simbody \\
\citet{Kargar2024}, ``Emerging Trends in Realistic Robotic Simulations'' & Gazebo, Webots, CoppeliaSim, MuJoCo, MORSE, PhysX, ODE, DART \\
\bottomrule
\end{tabular}
\end{table}

\section*{LLM-Assisted Discovery Sources}

Table~\ref{tab:llm-mention-sources} lists the auxiliary LLM-assisted discovery outputs archived with the source catalogue. These outputs were not treated as authoritative sources; they were used to surface additional candidate names that were then consolidated, filtered, and checked against public evidence before any package was selected for measurement.

\begin{table}[htbp]
\centering
\caption{LLM-assisted discovery sources archived with the source catalogue.}
\label{tab:llm-mention-sources}
\small
\singlespacing
\begin{tabular}{@{}p{4.6cm}p{8.2cm}@{}}
\toprule
\textbf{Source} & \textbf{Role in candidate discovery} \\
\midrule
GPT-generated robotics software list & Auxiliary broad-discovery list of open-source robotics software, simulators, planning libraries, physics engines, and related tooling. \\
Gemini-generated robotics software table & Auxiliary consolidated table of robotics frameworks, motion-planning tools, physics engines, and simulation platforms. \\
\bottomrule
\end{tabular}
\end{table}

\subsection*{LLM Query Details}

Both LLM-assisted outputs were generated on 2 November 2025 by issuing the following identical prompt, independently, to two assistants: GPT-5, accessed through the ChatGPT web interface with web browsing enabled, and Gemini~2.5~Pro, accessed through the Gemini web interface. Recording the model, prompt, and query date makes this discovery step reproducible in principle. The prompt instructed the models not to draw on prior conversational memory, so that each list reflects a fresh query rather than accumulated context.

\begin{quote}
\itshape
Based on authoritative sources, compile a list of approximately thirty robotics software packages related to motion planning, control, and simulation. Include both general-purpose frameworks and specialized toolkits, indicating their origin, primary domain focus, and level of usage or citation popularity. Do not use any prior conversational memory.
\end{quote}

The source catalogue with community-source URLs and archived LLM-assisted outputs, the raw captured text from the static public web lists, and the consolidated ranking with per-package mention counts are all preserved in the supplementary project materials that accompany this report.
