\chapter{Measurement Template}
\label{appendix:measurement-template}

This appendix reproduces the structure of the measurement template used in this report. The template is adapted from the Domain~X Appendix~B template~\citep{Smith2021} and is grouped into project metadata, the nine quality categories described in Section~\ref{sec:software-quality-attributes}, and repository metrics collected through the DomainX tooling. The full per-package responses are recorded in the project material \texttt{summary.csv}, with one row per question and one column per package. Each quality section ends with an overall impression score on a 1--10 scale.

\section*{Project Metadata}

\begin{itemize}
    \item Software name.
    \item URL.
    \item Affiliation: institution(s).
    \item Software purpose.
    \item Number of developers (commits-based count from repository logs).
    \item Funding source (unfunded, unclear, or funded with a string identifying the source).
    \item Initial release date.
    \item Last commit date.
    \item Status (alive, dead, or unclear; alive requires commits in the previous 18 months).
    \item License (GNU GPL, BSD, MIT, terms of use, trial, none, unclear, other).
    \item Supported platforms (Windows, Linux, macOS, Android, other).
    \item Software category (concept, public, or private).
    \item Development model (open source, freeware, commercial, or unclear).
    \item Number of publications mentioning the software.
    \item Source code URL.
    \item Programming languages.
    \item Whether performance was considered.
\end{itemize}

\section*{Installability}

\begin{itemize}
    \item Are installation instructions present?
    \item Are the instructions in one place (single document or web page)?
    \item Are the instructions linear (a single sequence of steps)?
    \item Are the instructions written assuming no dependencies are pre-installed?
    \item Are compatible operating system versions listed?
    \item Is the installation automated (makefile, script, installer, container)?
    \item If installation broke, was a descriptive error message displayed?
    \item Is there a specified way to validate the installation?
    \item Number of installation steps (including manual steps such as unzip).
    \item Operating system used for installation.
    \item Number of extra packages required before or during installation.
    \item Are required package versions listed?
    \item Are dependency installation instructions provided?
    \item Were there problems caused by uninstall (where applicable)?
    \item Overall installability impression (1--10).
\end{itemize}

\section*{Correctness and Verifiability}

\begin{itemize}
    \item Is there any reference to requirements specifications or theory manuals?
    \item What tools or techniques are used to build confidence in correctness (literate programming, automated testing, assertions, continuous integration, etc.)?
    \item Is there a getting-started tutorial?
    \item Are tutorial instructions linear?
    \item Does the tutorial provide an expected output?
    \item Does the tutorial output match the expected output?
    \item Are unit tests available?
    \item Is there evidence that performance was considered?
    \item Overall correctness and verifiability impression (1--10).
\end{itemize}

\section*{Surface Reliability}

\begin{itemize}
    \item Did the software break during installation?
    \item If installation broke, was a descriptive error message displayed?
    \item Did the software break during the initial tutorial test?
    \item If the tutorial test broke, was a descriptive error message displayed?
    \item If the tutorial test broke, was recovery possible?
    \item Overall surface reliability impression (1--10).
\end{itemize}

\section*{Surface Robustness}

\begin{itemize}
    \item Does the software handle unexpected or unanticipated input gracefully (data of the wrong type, malformed input, etc.)?
    \item For text input files, does the software handle modified line endings or other text-file edge cases?
    \item Overall surface robustness impression (1--10).
\end{itemize}

\section*{Surface Usability}

\begin{itemize}
    \item Is there a getting-started tutorial?
    \item Is there a user manual?
    \item Are expected user characteristics documented?
    \item What is the user support model (FAQ, user forum, email address for questions, etc.)?
    \item Overall surface usability impression (1--10).
\end{itemize}

\section*{Maintainability}

\begin{itemize}
    \item What is the current version number?
    \item Is there information on code review or contribution practices?
    \item Are non-code artifacts available, and which kinds?
    \item What issue tracking tool is in use?
    \item What percentage of identified issues are closed?
    \item What percentage of code consists of comments?
    \item Which version control system is in use?
    \item Overall maintainability impression (1--10).
\end{itemize}

\section*{Reusability}

\begin{itemize}
    \item How many code files are there?
    \item Is the API documented?
    \item Overall reusability impression (1--10).
\end{itemize}

\section*{Surface Understandability}

\begin{itemize}
    \item Is indentation and formatting style consistent?
    \item Is a coding standard explicitly identified?
    \item Are code identifiers consistent and meaningful?
    \item Are constants other than 0 and 1 hard-coded into the program?
    \item Are comments clear?
    \item Are parameters in the same order for all functions?
    \item Is the name or URL of any algorithm used mentioned?
    \item Is the code modularized?
    \item Overall surface understandability impression (1--10).
\end{itemize}

\section*{Visibility and Transparency}

\begin{itemize}
    \item Is the development process defined?
    \item Are there documents recording the development process and status?
    \item Is the development environment documented?
    \item Are release notes available?
    \item Overall visibility and transparency impression (1--10).
\end{itemize}

\section*{Repository Metrics}

Counts collected from the primary repository at the measurement date. For projects with multiple repositories, the choice of the primary repository is recorded as part of the measurement notes.

\begin{itemize}
    \item Number of text-based files.
    \item Number of binary files.
    \item Total lines in text-based files.
    \item Total lines added to text-based files.
    \item Total lines deleted from text-based files.
    \item Total number of commits.
    \item Commits per year for the last five years (or since project start if younger).
    \item Commits per month for the last twelve months.
    \item Code lines, comment lines, and blank lines in text-based files.
    \item Number of stars.
    \item Number of forks.
    \item Number of repository watchers.
    \item Number of open pull requests.
    \item Number of closed pull requests.
\end{itemize}

Each substantive question above corresponds to one row in \texttt{summary.csv}, with the per-package answer recorded in the column for that package. Including the supplementary entries (overall impression scores per category and the free-text additional-comments fields), the template contains approximately 108 fields per package.
