\chapter{Conclusions}
\label{ch:conclusions}

\section{Summary}
\label{sec:conclusion-summary}

This report assessed the state of the practice for robotics motion planning and
simulation (MPS) software by adapting the Domain~X methodology to a new domain.
Candidate packages were discovered from academic surveys, community-curated
lists, and auxiliary LLM-assisted discovery, consolidated into 295 unique
candidate names, ranked by mention count, and filtered by open-source
availability, repository-based measurability, and maintenance activity. The
resulting 28 packages, spanning full simulators, physics and dynamics engines,
planning libraries, manipulation frameworks, and middleware, were measured in
May 2026 against the Domain~X template: roughly 108 fields per package covering
project metadata, nine quality categories, and repository metrics. A
commonality analysis described the selected packages as a software family, and
the findings were compared with prior State-of-the-Practice studies of medical
imaging and Lattice Boltzmann method software. A developer-interview component
was designed but, pending ethics approval and recruitment, has not yet been
executed.

\section{Answers to Research Questions}
\label{sec:answers-research-questions}

\textbf{RQ1 (candidate packages).} Chapter~\ref{ch:methodology} and
Appendix~\ref{appendix:full-candidate-software-list} answer RQ1. Consolidating
the three discovery sources produced 295 unique candidate names. Filtering by
open-source availability, repository-based measurability, activity status, and
scope coherence reduced the leading candidates to the 28 measured packages;
notable exclusions include MATLAB/Simulink, Unity, NVIDIA Isaac Sim, and RaiSim
(closed core components), MRDS and USARSim (inactive), and GPMP2 and CHOMP
(single-algorithm scope).

\textbf{RQ2 (quality ranking).} Chapter~\ref{ch:measurement-results} answers
RQ2. Across the nine quality categories, the unweighted quality mean is led by
Gazebo (10.0), Drake (8.8), and a tied group of MoveIt!, OMPL, Pinocchio, and
ROS~2 (8.7); GraspIt! (5.0), Newton Dynamics (5.6), and Flightmare (6.1) rank
lowest, consistent with their inactive or thinly maintained status. Category
means are tightly bunched (7.3--7.8), so the rankings separate a strong top
tier and a weak bottom tier more clearly than they separate the middle of the
table. An elicited-weight AHP ranking remains future work; an illustrative
equal-weight computation reproduces the same leaders.

\textbf{RQ3 (repository artifacts).} Chapter~\ref{ch:comparison-research-software}
answers RQ3. Practical artifacts are near-universal in the MPS sample:
installation instructions, installation automation, tutorials, user manuals,
API documentation, issue tracking, and version control are present for all 28
packages, which compares favourably with the medical imaging and LBM samples.
The recurring gaps are formal artifacts: documented user characteristics (1 of
28) and explicit coding standards (14 of 28).

\textbf{RQ4 (tool usage).} The measured MPS projects are tightly connected to
modern open-source tooling: git-based version control and issue tracking are
universal, automated testing or CI evidence appears for 26 of 28 packages, and
documentation-generation tools are recorded for 12. Part of this uniformity is
a selection effect of the repository-based inclusion rule, but within the
measured set, tool adoption is stronger than in the LBM comparison point and at
least comparable with medical imaging.

\textbf{RQ5 (principles, processes, and methodologies).} Process evidence is
visible but artifact-bound: 25 of 28 packages expose a defined development
process, 27 publish release notes, and 26 provide contribution guidance. What
repositories do not reveal, design rationale, requirements assumptions, and
internal decision making, remains under-documented in MPS software, the same
broad pattern reported for other research software domains.

\textbf{RQ6--RQ9 (pain points and practices).} These questions require
developer-interview evidence and remain open. Chapter~\ref{ch:developer-interviews}
records the planned evidence base and reporting structure, and
Chapter~\ref{ch:recommendations} gives artifact-based recommendations that will
be revisited once interview findings are available.
% TODO (after interviews): replace this paragraph with summarized answers to
% RQ6--RQ9 based on the approved interview records.

\section{Key Findings}
\label{sec:key-findings}

Four findings stand out from the artifact evidence. First, the practical
engineering infrastructure of MPS software is strong: every measured package
can be installed from documented instructions with some form of automation, and
only two installations broke during measurement, both recoverably. Second, the
weaknesses are concentrated in formal and explanatory artifacts rather than in
tooling: user assumptions, coding standards, and design rationale are the
recurring gaps. Third, community popularity and measured quality are different
signals: Gazebo scores highest on quality while sitting in the middle of the
star ranking, and several highly starred packages place mid-table on quality,
so popularity indicators should not be used as quality proxies. Fourth,
activity status only loosely predicts measured quality: packages that go
inactive decay at different rates depending on the engineering capital they
accumulated while active, as the contrast between OpenRAVE and GraspIt!
illustrates.

\section{Future Work}
\label{sec:future-work}

The most immediate direction is the developer-interview component: completing
ethics approval, recruitment, and analysis would answer RQ6--RQ9 and connect
the measured artifact gaps to the pain points developers actually experience.
A second direction is an elicited-weight AHP ranking with sensitivity analysis,
which would turn the per-quality measurements into a defensible weighted
ordering. Longitudinal repeat measurements would track how the domain's
practices change over time, and the same methodology could be extended to
adjacent robotics software domains.

One further direction concerns the measurement process itself. The template
used in this study contains roughly one hundred fields per package
(Appendix~\ref{appendix:measurement-template}), and completing it by hand is
the most time-consuming part of the work. As an exploratory check, an AI coding
agent was asked to install, verify, and measure two of the selected physics
engines on its own, using only the official documentation and the measurement
template. The agent installed and verified both packages and filled the
template; its factual entries were tied to public sources, while its subjective
scores still needed human review. The trial is documented in
Appendix~\ref{appendix:ai-assisted-measurement}. With a person checking the
evidence, this kind of agent-assisted measurement could make it practical to
cover more packages, or to repeat the measurements over time, without straining
the time budget that constrains a manual study.
