\chapter{Introduction}
\label{ch:introduction}

Robotics software for motion planning and simulation enables researchers and engineers to model robot kinematics and dynamics, simulate sensor data, plan collision-free trajectories, and test control strategies before deploying to physical hardware. In practice, this software is central to experimental reproducibility and to the transition from simulation to physical deployment, so its engineering quality has direct consequences for both research and operational outcomes.

Developing this software takes considerable effort, and the engineering quality of the result matters. A simulator that produces incorrect trajectories, or a planner that crashes under realistic inputs, can compromise experimental results in ways that are not immediately obvious. Qualities such as correctness, reliability, and long-term sustainability are not guaranteed by domain expertise alone.

Software engineers have developed principles, methodologies, and tools to improve software quality, and practices such as version control, automated testing, and continuous integration are well established in professional contexts. There has historically been a gap, however, between these practices and what researchers actually do when developing scientific software. Scientists and engineers developing research software often acquire software engineering skills informally, from peers, through self-study, or by trial and error, rather than through formal training~\cite{Hannay2009}. Hannay et al.~\cite{Hannay2009} observe that many scientists show indifference to standard software engineering concepts, and Prabhu et al.~\cite{Prabhu2011} report that more than half of the scientists they surveyed did not use a debugger. Because much of the software studied in this report originates in research settings, these concerns apply directly to the domain.

Previous State-of-the-Practice studies have investigated this gap in several scientific software domains, including geographic information systems, mesh generation, seismology, statistical software for psychology, medical imaging software, and Lattice Boltzmann method software~\cite{smith2018state,smith2016state,Smith2018Seismology,smith2018statistical,Dong2021,Smith2024LBM}. This report applies the same methodology~\cite{Smith2021} to robotics software for motion planning and simulation (hereafter referred to as MPS software, where MPS stands for motion planning and simulation). The aim is to understand how MPS software holds up against established software engineering practices and how it compares with other research software domains. This means identifying where developers in this domain already meet those practices, where gaps remain, and whether other research software communities could draw on what MPS developers do. The report ranks the selected packages according to the defined quality attributes, but these rankings describe the current state of the practice and are not intended as a universal recommendation for any particular robotics task or context.

This chapter is organized as follows. Section~\ref{sec:research-questions} states the research questions. Section~\ref{sec:introduction-scope} defines the scope of the study. Section~\ref{sec:methodology-overview} provides a high-level overview of the methodology. Section~\ref{sec:contributions} describes the contributions of the report. Section~\ref{sec:organization} outlines the remaining chapters.

\section{Research Questions}
\label{sec:research-questions}

The following research questions guide this study. The abbreviation MPS (motion planning and simulation) is used throughout to refer to the studied software domain; the robotics context is implied.

\begin{enumerate}[RQ1.]
    \item What software packages are candidates for the State-of-the-Practice study of robotics software for motion planning and simulation?
    \item What is the ranking of the candidate packages from RQ1 with respect to the software qualities of installability, correctness and verifiability, reliability, robustness, usability, maintainability, reusability, understandability, and visibility and transparency?
    \item How do MPS projects compare to other research software domains with respect to the artifacts present in their repositories?
    \item How do MPS projects compare to other research software domains with respect to the use of tools?
    \item How do MPS projects compare to other research software domains with respect to the principles, processes, and methodologies used?
    \item What are the pain points for developers working on MPS software projects?
    \item How do the pain points for MPS developers compare to the pain points for developers in other research software domains?
    \item For MPS developers, what specific best practices address their pain points and software quality concerns?
    \item What research software development practices could potentially address the pain points identified in RQ6?
\end{enumerate}

RQ1 covers candidate discovery and selection. RQ2 addresses the quality measurement and ranking of selected packages. RQ3--RQ5 involve comparisons with prior State-of-the-Practice studies of other research software domains. RQ6--RQ9 address the developer perspective on pain points and best practices, and are conditional on approved interview data being available.

\section{Scope}
\label{sec:introduction-scope}

The working domain for this report is \emph{robotics software for motion planning and simulation}. This domain was established through discussion with the supervisor and the domain expert, Dr.~Matthew Giamou. It was initially discussed in relation to inverse kinematics software, but later project communication settled on the broader domain used here. The scope includes robotics simulators and closely related software used for planning, dynamics, physics, middleware, and robotic system development.

Several important and widely used tools fall outside the scope of this study. MATLAB and Simulink are major platforms in both robotics research and industry, but they are proprietary and do not permit the kind of public repository-based measurement that this study depends on. Dr.~Giamou confirmed that excluding MATLAB and Simulink does not misrepresent the domain. Unity, NVIDIA Isaac Sim, and RaiSim were excluded because their core simulation components are not open source, which prevents consistent repository mining across the study set. MRDS and USARSim were excluded because they are no longer actively maintained and do not represent current practice. GPMP2 and CHOMP, while relevant to optimization-based motion planning, are more narrowly scoped than the general-purpose tools in the candidate list and were excluded to maintain methodological consistency.

The study depends on evidence from public repositories, documentation, issue trackers, publications, and project metadata. Open-source availability and repository-based measurability are therefore essential selection criteria. A tool may be important in robotics practice and still fall outside the measured set if its core components are proprietary, if public repository evidence is unavailable, or if the project is no longer maintained enough to represent current practice.

The selected set is deliberately heterogeneous. Many entries are full robotics simulators, but ROS~2, MoveIt!, OMPL, and similar tools serve different roles. They remain in scope because motion planning and simulation workflows often depend on middleware, planning libraries, and supporting infrastructure. The comparison therefore treats the selected packages as related members of a domain, not as interchangeable products.

\section{Overview of the Methodology}
\label{sec:methodology-overview}

This report adapts the State-of-the-Practice methodology developed by Smith et al.~\cite{Smith2021}, which has previously been applied to domains including geographic information systems, mesh generation, seismology, statistical software for psychology, medical imaging, and Lattice Boltzmann method software~\cite{smith2018state,smith2016state,Smith2018Seismology,smith2018statistical,Dong2021,Smith2024LBM}. The method defines a measurable software domain, assembles a candidate package set from public sources (academic survey papers, community-curated lists, and auxiliary LLM-assisted discovery), and evaluates the selected packages using repository artifacts, documentation, releases, issue trackers, and project metadata. A commonality analysis examines the selected packages as a software family. If approved interview data becomes available, developer perspectives on pain points can supplement the artifact-based findings. Chapter~\ref{ch:methodology} gives the full procedures for candidate discovery, filtering, and measurement.

\section{Contributions}
\label{sec:contributions}

This report makes the following contributions:

\begin{enumerate}
    \item A scoped State-of-the-Practice study of robotics software for motion planning and simulation, applying the peer-reviewed methodology of Smith et al.~\cite{Smith2021} to a new domain.
    \item A transparent candidate discovery and selection process that combines academic sources, community-curated lists, and auxiliary LLM-assisted discovery, with explicit discussion of the role and limitations of each source type.
    \item A public-artifact measurement dataset for 28 selected software packages, covering repository evidence, documentation, releases, issue trackers, testing and continuous-integration evidence, and project metadata, organized across the nine quality categories defined in the Domain~X measurement template~\cite{Smith2021}.
    \item A commonality analysis identifying shared capabilities, variability points, and parameters of variation across the selected packages.
    \item A comparison of MPS software with findings from prior State-of-the-Practice studies in other research software domains~\cite{Smith2024LBM,Dong2021,Smith2018Seismology}, where supported by available evidence.
\end{enumerate}

These contributions are methodological and descriptive. The report characterizes existing software by measuring it against defined quality attributes; the resulting rankings describe the current state of the practice and do not prescribe a single best tool for all robotics tasks.

\section{Organization}
\label{sec:organization}

The remainder of this report is organized as follows:

\begin{itemize}
    \item \textbf{Chapter~\ref{ch:background}} introduces the robotics software domain and the motion planning and simulation context, reviews prior State-of-the-Practice studies, defines the software quality attributes used in this study, provides background on commonality analysis and the Analytic Hierarchy Process, and discusses the role and limitations of open-source repository-based measurement.
    \item \textbf{Chapter~\ref{ch:methodology}} describes the full methodology: domain selection, candidate discovery from multiple source types, filtering criteria, the measurement procedure using the Domain~X template~\cite{Smith2021}, commonality analysis procedure, and the planned interview component. This chapter addresses RQ1 and RQ2 at the procedural level.
    \item \textbf{Chapter~\ref{ch:selected-software-commonality}} presents the 28 selected packages as a software family and reports the commonality analysis, including shared capabilities, variability points, and parameters of variation. This chapter complements the package characterization from RQ1 by analyzing the selected software as a product family.
    \item \textbf{Chapter~\ref{ch:measurement-results}} reports the quality measurement results for the selected packages across the nine Domain~X quality categories, including repository metrics and overall observations. This chapter addresses RQ2.
    \item \textbf{Chapter~\ref{ch:comparison-research-software}} compares the MPS findings with results from prior State-of-the-Practice studies in other research software domains, examining similarities and differences with respect to repository artifacts, tool usage, and development practices. This chapter addresses RQ3--RQ5.
    \item \textbf{Chapter~\ref{ch:discussion}} interprets the measurement results, commonality analysis, and cross-domain comparison, and summarizes how the evidence answers the research questions.
    \item \textbf{Chapter~\ref{ch:developer-interviews}} describes the planned developer interview component, including the adapted Domain~X interview protocol and analysis approach. This component addresses RQ6--RQ9. Because no approved interview data is currently available, interview findings are reserved for a future study.
    \item \textbf{Chapter~\ref{ch:threats-to-validity}} discusses threats to the validity of the study, covering limitations of candidate selection, public-artifact measurement, source bias, and the interpretation of quality scores.
    \item \textbf{Chapter~\ref{ch:conclusions}} summarizes the study, provides high-level answers to the research questions, states the key findings, and identifies directions for future work.
\end{itemize}
