\chapter{Ranking Projects by Quality Measure}
\label{ch:measurement-results}

This chapter ranks the 28 selected software packages by measured engineering quality and compares that ranking with community-facing indicators of visibility and adoption. All data was collected during May 2026 and represents a snapshot of repository state, documentation, and project activity at that time. The quality measurements follow the Domain~X Appendix~B template and cover the nine quality categories described in Section~\ref{sec:software-quality-attributes}. Section~\ref{sec:repository-project-metadata} presents summary metadata. Sections~\ref{sec:installability-results} through~\ref{sec:visibility-transparency-results} report per-category results. Section~\ref{sec:repository-metrics-results} reports repository metrics. Section~\ref{sec:overall-observations} synthesizes the per-category findings and reports the quality-mean ranking. Section~\ref{sec:quality-vs-community} compares the quality-mean ranking with community indicators of visibility and adoption.

\section{Repository and Project Metadata}
\label{sec:repository-project-metadata}

Table~\ref{tab:summary-metadata} presents summary metadata for all 28 packages, including the primary development organization, license, initial release date, and primary programming languages. Platform support is summarized in the narrative following the table. The packages are listed alphabetically.

\begin{table}[H]
\centering
\caption{Summary metadata for the 28 selected software packages (May 2026).}
\label{tab:summary-metadata}
\small
\begin{tabular}{@{}p{2.5cm}p{3cm}p{1.8cm}p{1.4cm}p{1.8cm}@{}}
\toprule
\textbf{Software} & \textbf{Primary Affiliation} & \textbf{License} & \textbf{Initial Release} & \textbf{Primary Languages} \\
\midrule
AirSim & Microsoft Research & MIT & 2017 & C++, Python \\
Bullet/PyBullet & Community & zlib & 2006 & C++, Python \\
CARLA & CVC Barcelona, Intel & MIT & 2017 & C++, Python \\
CoppeliaSim & Coppelia Robotics & GPL + comm. & 2019 & C++, Python \\
DART & Georgia Tech & BSD & 2011 & C++, Python \\
Drake & TRI, MIT & BSD & 2010 & C++, Python \\
Flightmare & U. Zurich, ETH & MIT & 2020 & C++, Python \\
Gazebo & Open Robotics & Apache-2.0 & 2014 & C++, Python \\
GraspIt! & Columbia U. & GPL v3+ & 2010 & C++, Python \\
Habitat & Meta AI Research & MIT & 2019 & C++, Python \\
MORSE & LAAS-CNRS & BSD & 2009 & C, Python \\
MoveIt! & PickNik Robotics & BSD & 2011 & C++, Python \\
MRPT & U. Almer\'{i}a & BSD & 2010 & C++, Python \\
MuJoCo & Google DeepMind & Apache-2.0 & 2021 & C++, Python \\
Newton Dyn. & Independent & zlib & 2013 & C++, C \\
ODE & Independent & BSD/LGPL & 2001 & C \\
OMPL & Rice University & BSD & 2010 & C++, Python \\
OpenRAVE & CMU & LGPL + A2.0 & 2009 & C++, Python \\
OpenRTM-aist & AIST (Japan) & LGPL-2.1 & 2005 & C++, Python \\
OROCOS & KU Leuven & GPL + LGPL & 2007 & C++, Python \\
PhysX (OSS) & NVIDIA & BSD & 2022 & C++, Python \\
Pinocchio & Inria, LAAS-CNRS & BSD & 2014 & C++, Python \\
Proj.\ Chrono & U. Wisconsin, Parma & BSD & 2010 & C++, Python \\
ROS~2 & Open Robotics & Apache-2.0 & 2015 & C++, Python \\
Simbody & Stanford U. & Apache-2.0 & 2005 & C++, C \\
SOFA & Inria & LGPL-2.1 & 2005 & C++, Python \\
Webots & Cyberbotics & Apache-2.0 & 2018 & C++, Python \\
YARP & IIT (Italy) & BSD & 2006 & C++, Python \\
\bottomrule
\end{tabular}
\end{table}

All 28 packages have publicly accessible source repositories. Twenty-seven follow an open-source development model under standard permissive or copyleft licenses; CoppeliaSim is the exception, with a mixed model that combines a GPL-licensed open-source codebase with commercial and educational licenses for its packaged distribution. The most common licenses are BSD (11 packages), Apache~2.0 (5), and MIT (4). The packages span more than two decades of development history: ODE is the oldest, with its initial release in 2001, while PhysX's open-source SDK was released in 2022.

C++ is the dominant implementation language, used as a primary or major language by 26 of 28 packages. The two exceptions in the measurement records are MORSE (primary languages: C and Python) and ODE (primary language listed as unclear in the measurement template because the public source tree spans both C and C++ files without a single declared primary). Python is a primary or secondary language in 26 of 28 packages because of its growing role in robotics research and its use for scripting, bindings, and machine learning integration. Twenty-seven packages support Linux; the sole exception is PhysX, whose open-source SDK release listed only Windows as a target platform in the measurement data. Twenty-four packages support Windows and 23 support macOS.

As of the May 2026 measurement date, 24 of the 28 packages (86\%) were classified as alive under the Domain~X 18-month activity rule, which defines alive as the presence of any commit or release in the preceding 18 months. Four packages were classified as dead: OpenRAVE (last commit August 2024), MORSE (last commit March 2022), GraspIt! (last commit July 2020), and Flightmare (last commit May 2023). ODE is technically alive under the 18-month rule (last commit 2026-04-14) but its commit activity in the trailing twelve months is sparse (five commits across the year); the analysis that follows treats ODE as a low-activity alive package and discusses its limitations alongside the dead packages where appropriate.

The number of developers---defined as anyone who has made at least one accepted commit---ranges from 6 (CoppeliaSim, Flightmare) to 598 (MoveIt!). The median is approximately 128.5 (the mean of the 14th and 15th values, YARP at 125 and MRPT at 132). The total number of commits across all measured repositories exceeds 240,000. Drake leads with 34{,}786 commits, followed by SOFA (27{,}404) and Project Chrono (19{,}510). These repository-level metrics are snapshots and should not be interpreted as direct measures of project scale or quality without considering factors such as repository age, commit granularity, and whether development activity is concentrated in a single repository.

\section{Installability}
\label{sec:installability-results}

Installability measures the effort required to install and uninstall software in a specified environment. All installation measurements were performed on Linux-based virtual machines. Table~\ref{tab:installability} summarizes key installability indicators across the selected packages.

\begin{table}[H]
\centering
\caption{Key installability indicators across 28 packages.}
\label{tab:installability}
\begin{tabular}{@{}lrr@{}}
\toprule
\textbf{Indicator} & \textbf{Packages} & \textbf{\%} \\
\midrule
Installation instructions present & 28 & 100\% \\
Instructions in a single document or page & 26 & 93\% \\
Instructions are linear (single series of steps) & 26 & 93\% \\
Instructions assume no pre-installed dependencies & 22 & 79\% \\
Compatible OS versions listed & 17 & 61\% \\
Installation automation provided (script, makefile, etc.) & 28 & 100\% \\
Installation validation procedure available & 27 & 96\% \\
Dependency installation instructions provided & 24 & 86\% \\
Required package versions listed & 19 & 68\% \\
No problems caused by uninstall (where available) & 27 & 96\% \\
\bottomrule
\end{tabular}
\end{table}

All 28 packages provide installation instructions, and all provide some form of installation automation (makefiles, scripts, package managers, or container definitions). This is a strong result compared with prior State-of-the-Practice studies in other domains. For context, the Medical Imaging study~\cite{Dong2021} reported that only about half of MI packages were successfully installed.

The number of installation steps ranged from 1 to 11, with a median of 2. The number of extra software packages required before or during installation ranged from 0 to 10, with a median of 1. GraspIt! required the most manual effort (11 steps, 10 extra packages), consistent with its dead status and lack of recent maintenance. At the other extreme, several packages (Webots, MuJoCo, DART, Simbody, and others) required no extra packages and could be installed in a single step using package managers or pre-built binaries.

Installation broke for only 2 of the 28 packages during measurement (OpenRAVE and Newton Dynamics); in both cases the installation instance was recoverable. No package broke during initial tutorial testing.

The overall installability impression scores (on a 1--10 scale) range from 4 to 10, with a mean of 7.4. The top-scoring packages are Gazebo (10), Pinocchio (10), Bullet/PyBullet (9), MORSE (9), and OMPL (9). The lowest-scoring packages are ODE (4), PhysX (4), and GraspIt! (5). For ODE, the low score reflects outdated installation documentation (the official guide dates from 2006 and refers to GNU make, VC6 project files, and manual copying of library files). For PhysX, the low score reflects difficulty locating the correct build documentation.

\section{Correctness and Verifiability}
\label{sec:correctness-verifiability-results}

Correctness and verifiability assess whether a program matches its specification and the ease with which correctness can be checked. Table~\ref{tab:correctness-summary} summarizes key indicators.

\begin{table}[H]
\centering
\caption{Correctness and verifiability indicators.}
\label{tab:correctness-summary}
\begin{tabular}{@{}lrr@{}}
\toprule
\textbf{Indicator} & \textbf{Packages} & \textbf{\%} \\
\midrule
Reference to requirements specifications or theory manuals & 26 & 93\% \\
Getting-started tutorial present & 28 & 100\% \\
Tutorial instructions are linear & 27 & 96\% \\
Tutorial provides expected output & 26 & 93\% \\
Tutorial output matches expected output & 24 & 86\% \\
Unit tests available & 27 & 96\% \\
Evidence that performance was considered & 28 & 100\% \\
\bottomrule
\end{tabular}
\end{table}

All 28 packages use at least one technique to build confidence in correctness. The most common technique is automated testing, recorded for 26 of 28 packages (93\%). Assertions in code are used by 23 of 28 packages (82\%). Documentation generation tools such as Doxygen or Sphinx are used by 12 packages (43\%). Several packages combine multiple techniques: Drake, DART, Pinocchio, OMPL, Project Chrono, and Habitat use automated testing, assertions, and documentation generation together.

All 28 packages provide a getting-started tutorial, and 27 of 28 provide linear tutorial instructions. Twenty-six packages provide an expected tutorial output, and in 24 cases the measured output matched the expected output. The four cases where output did not match or was not applicable correspond to packages where the tutorial required environment-specific configuration not captured by the standard measurement procedure.

Unit tests are available for 27 of 28 packages. CoppeliaSim's unit test status is unclear because its test infrastructure is not publicly visible in the same way as the other packages, a consequence of its mixed commercial/open-source licensing model.

The overall correctness and verifiability impression scores range from 5 to 10, with a mean of 7.7. The top-scoring packages are Gazebo (10), Drake (10), ROS~2 (9), OMPL (9), Pinocchio (9), and MoveIt! (9). The lowest-scoring are GraspIt! (5) and Newton Dynamics (5).

\section{Surface Reliability}
\label{sec:surface-reliability-results}

Surface reliability assesses failure-free operation during the limited interactions performed during measurement: installation and initial tutorial use. These are surface checks, not comprehensive reliability experiments.

As noted in Section~\ref{sec:installability-results}, installation broke for 2 of 28 packages (OpenRAVE and Newton Dynamics). In both cases, a descriptive error message was displayed and the installation instance was recoverable. No package broke during initial tutorial testing.

The overall surface reliability impression scores range from 6 to 10, with a mean of 7.8. The top-scoring packages are Gazebo (10), Drake (10), ROS~2 (9), OMPL (9), Pinocchio (9), and MoveIt! (9). Three packages scored below 7: PhysX (6), GraspIt! (6), and Newton Dynamics (6).

The generally high surface reliability scores reflect the fact that most packages installed without incident and all tutorials ran without breakage. This is a positive signal for the domain, though it should be interpreted cautiously: the measurement procedure involves a shallow interaction (installation plus one tutorial) on a single operating system, and does not constitute a systematic reliability evaluation.

\section{Surface Robustness}
\label{sec:surface-robustness-results}

Surface robustness assesses whether software handles unexpected or unanticipated input reasonably, including edge cases such as malformed input files. All 28 packages were assessed as handling unexpected input reasonably (producing appropriate error messages or graceful degradation rather than crashes or undefined behaviour). Twenty-seven of 28 packages were assessed as handling newline conversion (replacing newlines with carriage returns and newlines in plain-text input files) gracefully.

The overall surface robustness impression scores range from 5 to 10, with a mean of 7.4. Gazebo scores highest (10), followed by Drake (9). The lowest-scoring package is GraspIt! (5). The narrow range of scores reflects the binary nature of the surface robustness checks in the Domain~X template: most packages either pass both checks or pass one of two, resulting in similar scores.

\section{Surface Usability}
\label{sec:surface-usability-results}

Surface usability assesses the presence and quality of user-facing documentation, tutorials, and support infrastructure. It is not a full usability evaluation; no user studies or task-completion experiments were conducted.

All 28 packages provide a getting-started tutorial and a user manual. This is a notably strong result. In contrast, the Medical Imaging study~\cite{Dong2021} found that only 22 of 29 MI packages (76\%) provided a user manual. Only one package in the current study (MuJoCo) documents expected user characteristics, which is consistent with findings from other domains where this practice is rare~\cite{Dong2021,Smith2024LBM}.

Table~\ref{tab:support-models} summarizes the user support models adopted by the study packages. Most packages use multiple support channels.

\begin{table}[H]
\centering
\caption{User support models across 28 packages.}
\label{tab:support-models}
\begin{tabular}{@{}lr@{}}
\toprule
\textbf{Support Channel} & \textbf{Packages Using} \\
\midrule
GitHub Issues & 28 \\
GitHub Discussions & 10 \\
Official forum or discussion board & 9 \\
Mailing list / Google Group & 8 \\
Discord server & 6 \\
Documentation website or wiki & 28 \\
Stack Exchange / Stack Overflow tag & 5 \\
ROS Answers integration & 4 \\
\bottomrule
\end{tabular}
\end{table}

GitHub Issues is the universal support channel, used by all 28 packages. Dedicated documentation websites are also universal. Community-oriented channels (forums, Discord servers, mailing lists) show more variation: 9 packages maintain official forums, 8 maintain mailing lists, and 6 maintain Discord servers. This variation partly reflects project age and governance model---newer packages and those stewarded by companies or large labs tend to favor Discord and GitHub Discussions, while older academic packages more commonly use mailing lists.

The overall surface usability impression scores range from 5 to 10, with a mean of 7.3. Gazebo (10), Drake (9), and MoveIt! (9) score highest; they have comprehensive documentation suites and multiple support channels. Newton Dynamics (5) and GraspIt! (5) score lowest.

\section{Maintainability}
\label{sec:maintainability-results}

Maintainability assesses the effort required to modify a software system or component. Table~\ref{tab:maintainability-summary} summarizes key maintainability indicators.

\begin{table}[H]
\centering
\caption{Maintainability indicators across 28 packages.}
\label{tab:maintainability-summary}
\begin{tabular}{@{}lrlr@{}}
\toprule
\textbf{Indicator} & \textbf{Count} & \textbf{\%} & \textbf{Notes} \\
\midrule
Version number documented & 28 & 100\% & --- \\
Code review / contribution guide & 26 & 93\% & --- \\
Non-code artifacts present & 28 & 100\% & All have config, docs, etc. \\
Issue tracking via git/GitHub & 28 & 100\% & --- \\
Version control via git/GitHub & 28 & 100\% & --- \\
\bottomrule
\end{tabular}
\end{table}

All 28 packages use git for version control and host their repositories on GitHub (with the exception of ODE, which uses BitBucket alongside a GitHub mirror). All 28 packages use GitHub Issues or the equivalent for issue tracking. This uniformity is partly a selection effect---the Domain~X methodology expresses a preference for GitHub-style repositories because they facilitate consistent measurement---but it also reflects the near-universal adoption of git and GitHub in the contemporary robotics software community.

The percentage of identified issues that are closed ranges from 34.6\% (Flightmare) to 100\% (DART, Newton Dynamics), with a mean of 79.8\% and a median of 82.3\%. A high closure rate can indicate active issue management, but it can also reflect a low rate of new issue reporting or a practice of closing stale issues aggressively. These percentages should be interpreted with caution.

The percentage of code that consists of comments ranges from 0.0\% (ROS~2, which is measured as zero code lines due to its multi-repository structure where the core communication libraries are compiled from source) to 27.98\% (OROCOS), with a mean of approximately 14.2\% excluding the ROS~2 outlier. OROCOS (27.98\%), OMPL (27.42\%), OpenRTM-aist (26.34\%), MoveIt! (26.12\%), and Simbody (25.27\%) lead in comment density. Webots (1.93\%), GraspIt! (2.05\%), and MORSE (2.67\%) have the lowest comment percentages. The Domain~X methodology treats 10\% as a threshold for the maintainability impression calculator.

The overall maintainability impression scores range from 4 to 10, with a mean of 7.5. Gazebo (10), ROS~2 (9), OMPL (9), Pinocchio (9), and MoveIt! (9) score highest. GraspIt! (4) and Flightmare (4) score lowest, weighed down by their dead or very low activity status and limited contribution infrastructure.

\section{Reusability}
\label{sec:reusability-results}

Reusability assesses the extent to which software components can be used in problem solutions other than the one for which they were originally developed. The Domain~X template measures reusability through two indicators: the number of code files (as a rough proxy for component granularity) and the presence of API documentation.

The number of code files varies widely across the selected packages, from 221 (OROCOS) to 7,823 (Newton Dynamics), with a median of approximately 1,500. All 28 packages provide API documentation. Where the measurement records identify documentation-generation tools, the most common tools are Doxygen and Sphinx.

The overall reusability impression scores range from 5 to 10, with a mean of 7.6. Gazebo (10), MuJoCo (9), OMPL (9), Pinocchio (9), ROS~2 (9), and MoveIt! (9) score highest. These packages are designed for integration into larger robotics systems or serve as the foundation that other tools build upon. MORSE (5) and GraspIt! (5) score lowest.

\section{Surface Understandability}
\label{sec:surface-understandability-results}

Surface understandability is assessed through a review of 10 randomly selected source files per package, examining code formatting, identifier quality, comment practices, and modularization. This is a shallow check, not a comprehensive code review.

Table~\ref{tab:understandability-summary} summarizes the surface understandability indicators.

\begin{table}[H]
\centering
\caption{Surface understandability indicators across 28 packages.}
\label{tab:understandability-summary}
\begin{tabular}{@{}lrr@{}}
\toprule
\textbf{Indicator} & \textbf{Packages} & \textbf{\%} \\
\midrule
Consistent indentation and formatting style & 28 & 100\% \\
Explicit identification of a coding standard & 14 & 50\% \\
Consistent, distinctive, and meaningful identifiers & 28 & 100\% \\
Hard-coded constants (other than 0 and 1) present & 28 & 100\% \\
Comments are clear and describe what is being done & 28 & 100\% \\
Name/URL of algorithms used is mentioned & 27 & 96\% \\
Code is modularized & 28 & 100\% \\
\bottomrule
\end{tabular}
\end{table}

All 28 packages exhibit consistent indentation and formatting, meaningful identifiers, clear comments, and modularized code organization. These are baseline indicators of professional software development practice, and their universal presence is a positive signal for the domain.

Fourteen of 28 packages (50\%) explicitly identify a coding standard. This is stronger than the MI comparison point, where code style guides were rare~\cite{Dong2021}, but it still leaves half of the measured MPS packages without an explicit coding-standard artifact. Packages that do identify a coding standard include Gazebo, Webots, Bullet/PyBullet, MuJoCo, ROS~2, Drake, OMPL, Pinocchio, MoveIt!, Habitat, Project Chrono, MRPT, YARP, and OpenRTM-aist.

The overall surface understandability impression scores range from 5 to 10, with a mean of 7.3. Gazebo (10), OMPL (9), and MoveIt! (9) score highest. GraspIt! (5) scores lowest.

\section{Visibility and Transparency}
\label{sec:visibility-transparency-results}

Visibility and transparency assess the extent to which a project's development process and status are visible to external observers. Table~\ref{tab:visibility-summary} summarizes the key indicators.

\begin{table}[H]
\centering
\caption{Visibility and transparency indicators.}
\label{tab:visibility-summary}
\begin{tabular}{@{}lrr@{}}
\toprule
\textbf{Indicator} & \textbf{Packages} & \textbf{\%} \\
\midrule
Development process defined (CI, contributing guide, etc.) & 25 & 89\% \\
Documents recording development process and status & 28 & 100\% \\
Development environment documented & 28 & 100\% \\
Release notes published & 27 & 96\% \\
\bottomrule
\end{tabular}
\end{table}

Twenty-five of 28 packages have a defined development process, typically manifested through CI/CD workflows, contributing guidelines, pull request templates, and issue templates. All 28 packages use GitHub Issues or similar tools to record development status. All 28 document their development environment (build dependencies, toolchain requirements, etc.). Twenty-seven of 28 publish release notes.

Three packages are recorded as unclear for the defined-development-process metric: CoppeliaSim, ODE, and OMPL. For CoppeliaSim, the mixed licensing model means parts of the development process are not publicly visible. For ODE, the unclear result is consistent with its low activity level and sparse process documentation. For OMPL, the result should be read narrowly as the outcome of this measurement item, not as a claim that the project lacks public development infrastructure in every respect.

The overall visibility and transparency impression scores range from 5 to 10, with a mean of 7.4. Gazebo (10), Drake (9), ROS~2 (9), OMPL (9), and MoveIt! (9) score highest. These packages publish thorough documentation of their development workflows. MORSE (5), Flightmare (5), Newton Dynamics (5), and GraspIt! (5) score lowest, consistent with their dead or low-activity status.

\section{Repository Metrics}
\label{sec:repository-metrics-results}

Table~\ref{tab:repository-metrics} presents key repository metrics for all 28 packages, including GitHub popularity indicators and codebase size measures. All data is from May 2026.

\begin{table}[H]
\centering
\caption{Repository metrics for the 28 selected packages (May 2026).}
\label{tab:repository-metrics}
\small
\begin{tabular}{@{}p{2.5cm}rrrrrr@{}}
\toprule
\textbf{Software} & \textbf{Stars} & \textbf{Forks} & \textbf{Commits} & \textbf{Code Lines} & \textbf{\% Comments} & \textbf{\% Issues Closed} \\
\midrule
AirSim & 18,159 & 4,889 & 3,562 & 134,331 & 10.3\% & 80.0\% \\
Bullet/PyBullet & 14,465 & 3,074 & 9,405 & 1,528,236 & 2.8\% & 87.3\% \\
CARLA & 13,941 & 4,559 & 6,713 & 138,801 & 9.4\% & 82.0\% \\
ROS~2 & 5,475 & 903 & 309 & --- & --- & 86.2\% \\
MuJoCo & 13,440 & 1,501 & 4,676 & 349,490 & 6.7\% & 91.7\% \\
Drake & 4,025 & 1,366 & 34,786 & 705,686 & 20.5\% & 90.0\% \\
Webots & 4,345 & 2,025 & 15,746 & 581,442 & 1.9\% & 87.9\% \\
Gazebo & 1,337 & 411 & 7,500 & 193,808 & 11.9\% & 56.7\% \\
PhysX (OSS) & 4,532 & 614 & 40$^\dagger$ & 1,030,524 & 17.3\% & 67.9\% \\
Pinocchio & 3,354 & 537 & 13,081 & 185,472 & 11.9\% & 92.3\% \\
Habitat & 3,662 & 530 & 1,664 & 143,150 & 7.7\% & 75.7\% \\
Proj.\ Chrono & 2,836 & 595 & 19,510 & 657,303 & 1.1\% & 96.8\% \\
Simbody & 2,521 & 492 & 5,041 & 259,447 & 25.3\% & 59.5\% \\
MRPT & 2,133 & 658 & 10,231 & 504,916 & 9.4\% & 95.4\% \\
OMPL & 2,047 & 690 & 5,341 & 105,532 & 27.4\% & 93.2\% \\
MoveIt! & 1,803 & 744 & 9,462 & 159,789 & 26.1\% & 80.5\% \\
Flightmare & 1,353 & 400 & 139 & 14,543 & 20.1\% & 34.6\% \\
SOFA & 1,194 & 352 & 27,404 & 349,961 & 4.5\% & 64.1\% \\
DART & 1,081 & 300 & 6,727 & 339,389 & 9.2\% & 100.0\% \\
CoppeliaSim & 146 & 46 & 1,206 & 292,460 & 5.1\% & 95.5\% \\
OROCOS & 879 & 443 & 1,229 & 18,063 & 28.0\% & 78.0\% \\
OpenRAVE & 801 & 355 & 10,375 & 623,306 & 19.2\% & 42.9\% \\
YARP & 590 & 214 & 19,020 & 843,982 & 15.1\% & 82.3\% \\
ODE & --- & --- & 2,119 & 147,055 & 18.8\% & --- \\
GraspIt! & 213 & 86 & 733 & 110,614 & 2.1\% & 56.6\% \\
OpenRTM-aist & 24 & 14 & 4,360 & 86,965 & 26.3\% & 81.4\% \\
MORSE & 367 & 156 & 4,356 & 141,241 & 2.7\% & 77.6\% \\
Newton Dyn. & 24 & 5 & 17,755 & 1,585,430 & 15.7\% & 100.0\% \\
\bottomrule
\end{tabular}
\end{table}

\noindent\textit{Notes:} ROS~2 code line and comment counts are not directly comparable because the ROS~2 organization spans hundreds of repositories. ODE is hosted on BitBucket; star, fork and issue-closure metrics are reported as ``\,---\,'' because the BitBucket APIs used during measurement did not return values for these fields under our access configuration, not because they are conceptually unavailable on that platform. The ``code lines'' column reports SCC-measured code lines in the primary measured repository. $^\dagger$ For PhysX, the commit count reflects only the public open-source SDK release on \texttt{NVIDIA-Omniverse/PhysX}; the broader internal NVIDIA codebase is not publicly accessible and is therefore not represented in this metric.

GitHub stars---a rough proxy for community attention---span four orders of magnitude, from 24 (OpenRTM-aist, Newton Dynamics) to 18,159 (AirSim). The median is approximately 2,300. The correlation between stars and measurement scores is not straightforward. AirSim leads in stars by a wide margin but does not appear among the top-scoring packages in most quality categories. Conversely, Gazebo scores highly across most quality categories but has relatively modest star counts (1,337), likely because its community engagement predates the current GitHub repository organization and because its user base interacts with it primarily through ROS rather than directly through its GitHub page.

The number of commits ranges from 139 (Flightmare, which is now inactive) to 34,786 (Drake). Commit activity in the last 12 months shows a similar range: Drake leads with sustained high monthly activity, while several dead packages (MORSE, GraspIt!, Flightmare) show zero or near-zero recent commits.

The percentage of code that is comments shows substantial variation, from 1.1\% (Project Chrono) to 28.0\% (OROCOS). The Domain~X maintainability calculator uses 10\% as a threshold: 20 of 28 packages (71\%) exceed this threshold. The packages below 10\% include some of the largest codebases in the study (Bullet/PyBullet at 2.8\%, SOFA at 4.5\%, Project Chrono at 1.1\%), suggesting that codebase size alone does not determine comment density.

\section{Overall Observations}
\label{sec:overall-observations}

Table~\ref{tab:overall-impressions} summarizes the overall impression scores across all nine quality categories for each package, along with a simple unweighted per-package mean. The scores reported here are the raw 1--10 impression values assigned during measurement using the Domain~X Appendix~C calculator. They are \emph{not} the definitive ranking of the study: that role is reserved for the AHP-based analysis described in Chapter~\ref{ch:methodology}, which accounts for differential importance weights across qualities.

\begin{table}[H]
\centering
\caption{Overall impression scores by quality category (1--10 scale) with an unweighted per-package mean. I = Installability, C = Correctness, SR = Surface Reliability, SRo = Surface Robustness, SU = Surface Usability, M = Maintainability, Re = Reusability, SUnd = Surface Understandability, VT = Visibility/Transparency. The Mean column is the simple arithmetic mean of the nine category scores and is intentionally unweighted; it is not a substitute for the AHP-based ranking, which would apply differential importance weights across qualities.}
\label{tab:overall-impressions}
\small
\begin{tabular}{@{}p{2.5cm}cccccccccc@{}}
\toprule
\textbf{Software} & \textbf{I} & \textbf{C} & \textbf{SR} & \textbf{SRo} & \textbf{SU} & \textbf{M} & \textbf{Re} & \textbf{SUnd} & \textbf{VT} & \textbf{Mean} \\
\midrule
AirSim & 8 & 8 & 8 & 8 & 8 & 8 & 8 & 7 & 7 & 7.8 \\
Bullet/PyBullet & 9 & 7 & 8 & 7 & 7 & 7 & 8 & 7 & 7 & 7.4 \\
CARLA & 7 & 8 & 8 & 7 & 8 & 8 & 8 & 7 & 8 & 7.7 \\
CoppeliaSim & 8 & 8 & 8 & 7 & 7 & 7 & 8 & 7 & 7 & 7.4 \\
DART & 8 & 8 & 8 & 7 & 7 & 8 & 8 & 7 & 7 & 7.6 \\
Drake & 7 & 10 & 10 & 9 & 9 & 9 & 8 & 9 & 9 & 8.9 \\
Flightmare & 7 & 6 & 7 & 8 & 6 & 4 & 6 & 6 & 5 & 6.1 \\
Gazebo & 10 & 10 & 10 & 10 & 10 & 10 & 10 & 10 & 10 & 10.0 \\
GraspIt! & 5 & 5 & 6 & 5 & 5 & 4 & 5 & 5 & 5 & 5.0 \\
Habitat & 8 & 8 & 8 & 8 & 8 & 8 & 8 & 8 & 8 & 8.0 \\
MORSE & 9 & 8 & 8 & 8 & 8 & 8 & 5 & 6 & 5 & 7.2 \\
MoveIt! & 7 & 9 & 9 & 8 & 9 & 9 & 9 & 9 & 9 & 8.7 \\
MRPT & 8 & 8 & 7 & 8 & 7 & 8 & 8 & 8 & 8 & 7.8 \\
MuJoCo & 7 & 7 & 8 & 8 & 8 & 8 & 9 & 8 & 8 & 7.9 \\
Newton Dyn. & 6 & 5 & 6 & 6 & 5 & 5 & 6 & 6 & 5 & 5.6 \\
ODE & 4 & 6 & 7 & 7 & 6 & 6 & 7 & 7 & 7 & 6.3 \\
OMPL & 9 & 9 & 9 & 8 & 8 & 9 & 9 & 9 & 9 & 8.8 \\
OpenRAVE & 8 & 8 & 8 & 7 & 8 & 8 & 7 & 8 & 7 & 7.7 \\
OpenRTM-aist & 7 & 8 & 7 & 7 & 6 & 7 & 6 & 7 & 7 & 6.9 \\
OROCOS & 7 & 8 & 7 & 6 & 6 & 6 & 7 & 6 & 6 & 6.6 \\
PhysX (OSS) & 4 & 6 & 6 & 6 & 7 & 8 & 8 & 8 & 8 & 6.8 \\
Pinocchio & 10 & 9 & 9 & 8 & 8 & 9 & 9 & 8 & 8 & 8.7 \\
Proj.\ Chrono & 7 & 8 & 8 & 7 & 7 & 8 & 7 & 7 & 8 & 7.4 \\
ROS~2 & 8 & 9 & 9 & 8 & 9 & 9 & 9 & 8 & 9 & 8.7 \\
Simbody & 7 & 7 & 7 & 7 & 7 & 7 & 7 & 7 & 7 & 7.0 \\
SOFA & 6 & 6 & 8 & 7 & 6 & 7 & 7 & 7 & 7 & 6.8 \\
Webots & 8 & 8 & 8 & 7 & 8 & 8 & 8 & 7 & 8 & 7.8 \\
YARP & 7 & 8 & 7 & 7 & 7 & 8 & 7 & 7 & 7 & 7.2 \\
\bottomrule
\end{tabular}
\end{table}

Several patterns emerge from the measurement results:

\textbf{High performers across multiple qualities.} Gazebo scores 10 in all nine categories, making it the most consistently high-rated package in the study. Drake and MoveIt! score 9 or 10 across most categories. ROS~2, OMPL, and Pinocchio also show strong scores across the board. These packages share several characteristics: they are stewarded by established organizations or research groups, have active contributor communities, maintain comprehensive documentation, use automated testing and CI/CD, and publish clear contribution guidelines and release notes.

\textbf{Dead or inactive packages score lower across the board.} GraspIt! (last commit July 2020), Flightmare (last commit May 2023), and MORSE (last commit March 2022) appear in the bottom tier for most quality categories. This is expected: packages that are no longer maintained tend to accumulate outdated documentation, broken installation procedures, and unresolved issues. Newton Dynamics, while technically alive (last commit May 2026), scores in the bottom tier, reflecting its status as a primarily individual-maintained project with limited documentation infrastructure.

\textbf{Installability is generally strong.} All 28 packages provide installation instructions and automation, and only 2 broke during installation. This compares favorably with prior State-of-the-Practice studies. The widespread adoption of package managers (APT, pip, Conda), containerization (Docker), and standardized build systems (CMake) in the robotics community likely contributes to this result.

\textbf{Documentation is universally present but varies in depth.} All packages provide tutorials, user manuals, and API documentation. However, only one package documents expected user characteristics, and only 14 explicitly identify a coding standard. These gaps are consistent with findings in medical imaging and LBM software~\cite{Dong2021,Smith2024LBM} and reflect broader patterns in research software development.

\textbf{Repository activity is concentrated in a subset of packages.} A small number of packages (Drake, SOFA, Project Chrono, YARP, Pinocchio) account for a disproportionate share of total commits. This reflects sustained multi-year development by well-resourced teams. At the other extreme, several packages show minimal activity in the last 12 months.

\textbf{The relationship between community popularity and measurement scores is not linear.} Section~\ref{sec:quality-vs-community} reports the explicit ranking comparison; popularity metrics are cumulative and backward-looking, while measurement scores reflect current documentation and process quality.

These observations are descriptive and are based on the specific measurement procedure, time frame, and interpretation rules described in Chapter~\ref{ch:methodology}. They are not claims that any single package is universally superior. Chapters~\ref{ch:comparison-research-software} and~\ref{ch:discussion} interpret these results in the context of the research questions.

\section{Comparison with Community Indicators}
\label{sec:quality-vs-community}

The Domain~X quality mean reported in Section~\ref{sec:overall-observations} reflects the per-package state of documentation, testing, release management, and related software engineering practices in May 2026. Community indicators such as GitHub stars, forks, and recent commit activity capture a different signal: cumulative visibility, user adoption, and word-of-mouth attention built up over time. The two need not agree. This section presents both rankings side by side, using the per-package quality mean from Table~\ref{tab:overall-impressions} and the star count from Table~\ref{tab:repository-metrics}, and identifies where the indicators converge and where they diverge.

\begin{table}[H]
\centering
\caption{Quality-mean rank versus GitHub stars rank for the 28 selected packages. $\Delta = S - Q$: positive values indicate a higher quality rank than stars rank (under-recognized for its quality); negative values indicate a higher stars rank than quality rank (popular without proportional quality scores). Ties receive the same rank. ODE is hosted on BitBucket and is excluded from the stars rank, consistent with Table~\ref{tab:repository-metrics}.}
\label{tab:quality-vs-community}
\small
\begin{tabular}{@{}p{2.5cm}cccrr@{}}
\toprule
\textbf{Software} & \textbf{Q-Mean} & \textbf{Q-Rank} & \textbf{Stars} & \textbf{S-Rank} & \textbf{$\Delta$} \\
\midrule
Gazebo           & 10.0 & 1  & 1,337  & 17 & $+16$ \\
Drake            & 8.9  & 2  & 4,025  & 8  & $+6$  \\
OMPL             & 8.8  & 3  & 2,047  & 14 & $+11$ \\
MoveIt!          & 8.7  & 4  & 1,803  & 15 & $+11$ \\
Pinocchio        & 8.7  & 4  & 3,354  & 10 & $+6$  \\
ROS~2            & 8.7  & 4  & 5,475  & 5  & $+1$  \\
Habitat          & 8.0  & 7  & 3,662  & 9  & $+2$  \\
MuJoCo           & 7.9  & 8  & 13,440 & 4  & $-4$  \\
AirSim           & 7.8  & 9  & 18,159 & 1  & $-8$  \\
MRPT             & 7.8  & 9  & 2,133  & 13 & $+4$  \\
Webots           & 7.8  & 9  & 4,345  & 7  & $-2$  \\
CARLA            & 7.7  & 12 & 13,941 & 3  & $-9$  \\
OpenRAVE         & 7.7  & 12 & 801    & 21 & $+9$  \\
DART             & 7.6  & 14 & 1,081  & 19 & $+5$  \\
Bullet/PyBullet  & 7.4  & 15 & 14,465 & 2  & $-13$ \\
CoppeliaSim      & 7.4  & 15 & 146    & 25 & $+10$ \\
Proj.\ Chrono    & 7.4  & 15 & 2,836  & 11 & $-4$  \\
MORSE            & 7.2  & 18 & 367    & 23 & $+5$  \\
YARP             & 7.2  & 18 & 590    & 22 & $+4$  \\
Simbody          & 7.0  & 20 & 2,521  & 12 & $-8$  \\
OpenRTM-aist     & 6.9  & 21 & 24     & 26 & $+5$  \\
PhysX (OSS)      & 6.8  & 22 & 4,532  & 6  & $-16$ \\
SOFA             & 6.8  & 22 & 1,194  & 18 & $-4$  \\
OROCOS           & 6.6  & 24 & 879    & 20 & $-4$  \\
ODE              & 6.3  & 25 & ---    & ---& ---   \\
Flightmare       & 6.1  & 26 & 1,353  & 16 & $-10$ \\
Newton Dyn.      & 5.6  & 27 & 24     & 26 & $-1$  \\
GraspIt!         & 5.0  & 28 & 213    & 24 & $-4$  \\
\bottomrule
\end{tabular}
\end{table}

The 28 packages fall into four broad patterns. \textbf{Rough agreement at the top.} Drake ($\Delta = +6$), ROS~2 ($\Delta = +1$), Pinocchio ($\Delta = +6$), and Habitat ($\Delta = +2$) appear in the top tier on both indicators; well-resourced projects with sustained development tend to combine strong engineering practice with substantial visibility. \textbf{High quality, low visibility.} Gazebo ($\Delta = +16$), OMPL ($\Delta = +11$), and MoveIt! ($\Delta = +11$) score at or near the top on quality but sit well below their quality rank on stars. The Gazebo gap is partially explained at the end of Section~\ref{sec:repository-metrics-results}: Gazebo's user base interacts with it primarily through ROS rather than its GitHub repository directly, so star counts undercount its reach. OMPL and MoveIt! are libraries embedded inside larger workflows and may attract attention as dependencies rather than as standalone projects. \textbf{High visibility, mid-pack quality.} AirSim ($\Delta = -8$), Bullet/PyBullet ($\Delta = -13$), CARLA ($\Delta = -9$), and PhysX (OSS) ($\Delta = -16$) sit at or near the top on stars but in the middle or lower tier on the quality mean. These projects share characteristics: broad user bases drawn from adjacent fields (gaming, autonomous-driving research, real-time physics), corporate or game-engine origins, and uneven research-software quality practice. \textbf{Consistently low.} GraspIt! ($\Delta = -4$), Newton Dynamics ($\Delta = -1$), and OpenRTM-aist ($\Delta = +5$) cluster near the bottom on both indicators; inactive or narrowly used projects tend to accumulate gaps in documentation, releases, and visibility together.

These patterns are consistent with the finding from the Medical Imaging State-of-the-Practice study that GitHub popularity and measurement-based rankings do not align in a simple way~\cite{Dong2021}. They support the broader argument that community-facing indicators of visibility and adoption are not interchangeable with measured software engineering quality; treating stars as a quality proxy obscures meaningful differences across packages.
