\chapter{Methodology}
\label{ch:methodology}

This chapter presents the study methodology and addresses RQ1 (software candidate identification) by documenting how the candidate set was assembled. The methodology follows \citet{Smith2021}, adapted to the MPS domain, in the following steps:

\begin{enumerate}
    \item \textbf{Select the research domain} (Section~\ref{sec:domain-selection}).
    \item \textbf{Engage a domain expert} who participates in the assessment and later reviews the candidate list (Section~\ref{sec:domain-expert-interaction}).
    \item \textbf{Identify candidate software} from academic survey papers, community-curated lists, and LLM-assisted lists (Section~\ref{sec:candidate-software-discovery}).
    \item \textbf{Rank the candidates} by mention count (Section~\ref{sec:mention-count-ranking}).
    \item \textbf{Filter the ranked list} to the selected set of 28 packages (Sections~\ref{sec:filtering-criteria} and~\ref{sec:final-software-set}).
    \item \textbf{Measure each package} with the Domain~X template and rank the results with AHP (Section~\ref{sec:measurement-procedure}).
    \item \textbf{Analyze commonalities} across the selected packages (Section~\ref{sec:commonality-analysis-procedure}).
    \item \textbf{Interview developers} about their development practices and pain points (Section~\ref{sec:interview-component}).
\end{enumerate}

\section{Domain Selection}
\label{sec:domain-selection}

The first step of the Domain~X methodology is to identify a suitable research software domain. To be applicable for the methodology, the chosen domain must meet four criteria~\citep{Smith2021}: (i)~it must have a well-defined and stable theoretical underpinning, ideally supported by standard textbooks; (ii)~there must be a community of researchers and practitioners studying the domain; (iii)~the software packages in the domain must include open-source options; and (iv)~a preliminary search should suggest that there are close to 30 candidate software packages that qualify as research software.

The selected domain is \emph{MPS (motion planning and simulation)}. Motion planning and simulation are established subfields of robotics with well-defined theoretical foundations~\citep{lavalle2006planning,choset2005principles}. There is an active international research community, regular conferences (such as ICRA, IROS, and RSS), and a substantial body of published work. A preliminary search confirmed the existence of numerous open-source software packages across the MPS software roles defined in Section~\ref{sec:robotics-software-background}.

The domain boundary follows the MPS scope defined in Chapter~\ref{ch:background}, Section~\ref{sec:robotics-software-background}, rather than a simulator-only scope.

Earlier project discussions considered a narrower domain focused on inverse kinematics software. After consultation with the supervisor and domain expert, the scope was broadened to the MPS domain, which provides a richer set of candidate packages and better reflects the interconnected nature of robotics software ecosystems.

\section{Interaction with the Domain Expert}
\label{sec:domain-expert-interaction}

The Domain~X methodology emphasizes that a domain expert is an essential member of the assessment team. Pitfalls exist if non-experts attempt to produce an authoritative list of software or definitively rank packages without domain knowledge. Non-experts must rely on information available online, which has two known drawbacks: (i)~online resources may contain false or inaccurate information; and (ii)~online resources may omit relevant information that is so ingrained with experts that nobody thinks to record it explicitly~\citep{Smith2021}.

The expert need not be a software developer, only a knowledgeable user of domain software, and the methodology is designed to minimize the demand on their time~\citep{Smith2024MI}.

For this report, the domain expert is Dr.~Matthew Giamou of the Department of Computing and Software at McMaster University, whose research includes robotics, motion planning, and state estimation. His review of the candidate list and the filtering outcome is reported in Section~\ref{sec:final-software-set}, after the preliminary list has been assembled.

\section{Candidate Software Discovery}
\label{sec:candidate-software-discovery}

Candidate software packages were identified from three types of sources: academic survey papers (Section~\ref{subsec:academic-sources}), community-curated software lists (Section~\ref{subsec:community-curated-sources}), and LLM-generated software lists (Section~\ref{subsec:llm-assisted-sources}). Using multiple source types reduces dependence on any single search method and provides broader coverage of the MPS domain. The collected names were consolidated to merge equivalent entries (for example, ``V-REP'' and ``CoppeliaSim'' refer to the same package), and the consolidated list was used as input to the mention-count ranking described in Section~\ref{sec:mention-count-ranking}.

\subsection{Academic Survey Papers}
\label{subsec:academic-sources}

Academic sources identify software packages that appear in research discussions of robotics simulators, simulation tools, and related challenges. These sources included survey papers and comparative studies, such as \citet{Harris2011}, \citet{collins2021simulators}, \citet{Korber2021}, and \citet{Farley2022}, that discuss packages such as Gazebo, Webots, V-REP/CoppeliaSim, MuJoCo, and PyBullet in the context of robotics research. The full list of academic sources used for mention counting appears in Appendix~\ref{appendix:mention-count-source-list}.

Academic sources have known limitations as the sole basis for candidate discovery. Survey papers tend to emphasize packages relevant to a specific research question, robot type, or time period; some active tools may not appear in any of the surveys consulted, while older tools sometimes remain visible because they were prominent when a survey was written.

\subsection{Community-Curated Sources}
\label{subsec:community-curated-sources}

Community-curated sources broaden the view of robotics software practice beyond what appears in academic surveys. The source catalogue included GitHub Topic pages, ``awesome'' lists (curated collections such as \texttt{awesome-robotics-libraries}, \texttt{awesome-robotic-tooling}, and \texttt{awesome-motion-planning}), best-of lists such as \texttt{best-of-robot-simulators}, and other community-maintained aggregations. Sources covered topics including motion planning, robotics simulation, robotics libraries, ROS~2 ecosystem tools, multibody dynamics simulation, and general robotic tooling. The GitHub Topic pages were manually inspected as dynamic community aggregations, while the static curated-list pages were preserved as copied raw web material where available. Appendix~\ref{appendix:mention-count-source-list} lists the community-curated sources individually.

Community-curated sources vary in maintenance quality, inclusion criteria, and update frequency, so they were used as discovery aids rather than as indicators of software quality.

\subsection{LLM-Assisted Sources}
\label{subsec:llm-assisted-sources}

LLM-generated software lists were used as an auxiliary source for candidate discovery. This modifies the original Domain~X methodology, which prescribes search engine queries and domain expert consultation as the primary discovery methods. The modification is recorded explicitly so that the discovery step remains transparent and auditable. The model, prompt, and query date for each LLM-generated list are recorded in Appendix~\ref{appendix:mention-count-source-list}.

The reason for including LLM-assisted discovery is practical: a large language model acts as a broad aggregator over existing public information, analogous in effect to a wide search over indexed web content. LLM output can also be incomplete, outdated, or incorrect, so names obtained from LLM-generated lists were consolidated, filtered, and verified against public evidence before any claims in this report were based on them.

\section{Mention Count Ranking}
\label{sec:mention-count-ranking}

After consolidating equivalent names across the collected sources, each unique software package was counted once per source to produce a mention count. A package that appeared in five academic papers and three community-curated lists would receive a mention count of eight. The consolidated ranking was recorded during the discovery phase and is a transparent input to the filtering process.

Consolidation produced 295 unique candidate names. The top entries in the consolidated ranking include Gazebo (22 mentions), Webots (16), Bullet/PyBullet (16), CoppeliaSim/V-REP (13), MuJoCo (12), ODE (10), ROS~2 (10), and OpenRAVE (9). Appendix~\ref{appendix:full-candidate-software-list} records every candidate with at least four mentions, together with its mention count and filtering outcome, and Appendix~\ref{appendix:mention-count-source-list} lists the counted sources. The full ranked list of all 295 names, including packages that did not survive the filtering step described in Section~\ref{sec:filtering-criteria}, is preserved in the supplementary project materials that accompany this report (a consolidated ranking spreadsheet and a companion source catalogue).\footnote{The supplementary project materials are available in the project repository at \suppurl.}

Mention count is a source-appearance signal and must be interpreted accordingly. It does not measure software quality, user adoption, research impact, maintainability, or the overall suitability of any particular software package. A package may appear often because it is historically established or broadly useful, while a newer or more specialized package may be under-represented even when it is technically strong.

\section{Filtering Criteria}
\label{sec:filtering-criteria}

The ranked candidate list was filtered to identify packages suitable for repository-based measurement of the state of the practice. In the Domain~X methodology, the initial filters are scope, usage, and age, applied in priority order until the list reaches a manageable size of approximately 30 packages~\citep{Smith2021}. In this report, these filters are implemented through three criteria: open-source availability, repository-based measurability, and activity and maintenance status.

\subsection{Open Source Availability}
\label{subsec:open-source-availability}

Open-source availability was required because the study depends on public software artifacts for measurement. The Domain~X candidate criteria require that source code be viewable and express a preference for GitHub-style repositories from which empirical measures can be collected~\citep{Smith2021}. Source code, repository history, issue trackers, documentation, and release information provide the evidence base for the measurements reported in Chapter~\ref{ch:measurement-results}.

This criterion led to the exclusion of several prominent robotics tools whose core components are not open source:

\begin{itemize}
    \item \textbf{MATLAB and Simulink} are major platforms in both robotics research and industry, widely used for algorithm development, control design, and simulation.
    \item \textbf{Unity} is a widely used game engine with robotics simulation capabilities.
    \item \textbf{NVIDIA Isaac Sim} builds on NVIDIA Omniverse and provides advanced robotics simulation features.
    \item \textbf{RaiSim} is a high-performance physics simulator for robotics.
\end{itemize}

The exclusions above are methodological constraints, not judgements about the quality or importance of the excluded tools; the resulting coverage limit is discussed in Chapter~\ref{ch:threats-to-validity}.

\subsection{Repository-Based Measurability}
\label{subsec:repository-based-measurability}

Repository-based measurability requires that a package have sufficient public evidence to support the measurement process described in Section~\ref{sec:measurement-procedure}. Relevant evidence includes source repositories, commit history, issue data, documentation, release information, build or installation instructions, test artifacts, and project metadata. Packages without sufficient public evidence cannot be measured consistently and were excluded.

This criterion also governs how multi-repository projects are handled. Some robotics software systems are distributed across multiple repositories or GitHub organizations. In such cases, the repositories most representative of the core software were selected for measurement, and this choice was documented as part of the measurement record. The same principle applies to projects that have changed names, migrated repositories, or spawned successor projects.

\subsection{Activity and Maintenance Status}
\label{subsec:activity-maintenance-status}

Activity and maintenance status were considered because the study focuses on contemporary software practice. In the Domain~X methodology, age is one of the initial filtering criteria, measured by the last date when a change was made, with exceptions for older packages that remain highly recommended and currently in use~\citep{Smith2021}. Tools that are no longer maintained provide weaker evidence about current development practices.

The primary activity indicator used in this report is the Domain~X 18-month repository-activity flag: a package receives the value \emph{alive} if its measured repository has received any commit or version release within the 18 months preceding the measurement date, and \emph{dead} otherwise. The flag records repository events; it does not by itself establish that substantive software development or project stewardship is continuing. For example, documentation-only commits can produce an alive flag even when a project's maintainers have announced that development has ended. The flag was therefore used as a weighted signal rather than as an automatic cutoff: a package was set aside on activity grounds only when prolonged inactivity coincided with the tool no longer being prominent or having been superseded by an actively maintained alternative. On that basis, the following packages were excluded:

\begin{itemize}
    \item \textbf{MRDS} (Microsoft Robotics Developer Studio) has not received updates since 2012 and is no longer supported by Microsoft.
    \item \textbf{USARSim} (Unified System for Automation and Robot Simulation) is no longer actively maintained and does not represent current practice.
\end{itemize}

Four packages (OpenRAVE, MORSE, GraspIt!, and Flightmare) received a dead repository-activity flag at the measurement date but remain prominent and widely referenced, and so were retained in the measured set. Their flags are reported as measured outcomes in Chapter~\ref{ch:measurement-results} (Section~\ref{sec:repository-project-metadata}), where they are interpreted together with project-specific maintenance evidence rather than as stand-alone claims about project viability.

\subsection{Scope and Coverage Filters}
\label{subsec:scope-coverage-filters}

Two algorithm-level reference implementations were excluded to keep the selected packages methodologically coherent rather than for quality reasons. \textbf{GPMP2} (Gaussian Process Motion Planner 2) and \textbf{CHOMP} (Covariant Hamiltonian Optimization for Motion Planning) are optimization-based motion planning algorithms with public reference implementations. They are relevant to motion planning research but are narrower in scope than the general-purpose simulators, physics engines, planning frameworks, and middleware tools that dominate the candidate list. Including them alongside Gazebo, MoveIt!, or ROS~2 would force a comparison between full software stacks and single-algorithm libraries, which would distort the commonality analysis and the cross-package measurements.

Filtering decisions use a measurement-time snapshot of repository activity and release status; these signals can change afterward, so the measurement date is recorded with the results (Chapter~\ref{ch:measurement-results}).

\section{Final Software Set}
\label{sec:final-software-set}

After ranking by mention count and applying the filtering criteria described above, the selected set consists of 28 software packages. Appendix~\ref{appendix:full-candidate-software-list} records the disposition of the leading candidates, and Appendix~\ref{appendix:excluded-software-packages} details the excluded packages. Figure~\ref{fig:candidate-selection} summarizes the full selection pipeline, from the three discovery sources through the mention-count ranking and the filtering criteria to the final set. Table~\ref{tab:final-software-set} lists the packages alphabetically with their primary roles in the robotics software ecosystem.

The domain expert, Dr.~Matthew Giamou, reviewed the candidate list and the filtering outcome. He confirmed that the list was reasonable and that most of the shortlisted packages were familiar to him, and he noted that ROS~2, MoveIt!, and OMPL are not full simulators but fit within a coherent narrative when their roles are explained carefully. He agreed that excluding MATLAB/Simulink, Unity, NVIDIA Isaac Sim, and RaiSim on open-source grounds (Section~\ref{subsec:open-source-availability}) would not misrepresent the MPS domain; MRDS and USARSim were filtered on activity grounds (Section~\ref{subsec:activity-maintenance-status}) rather than by a separate expert decision.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
    font=\small,
    box/.style={draw, rounded corners, align=center, inner sep=5pt, minimum height=0.85cm},
    src/.style={box, fill=blue!5, text width=2.9cm},
    proc/.style={box, fill=black!4, text width=8.4cm},
    final/.style={box, fill=blue!12, text width=5.2cm, font=\small\bfseries},
    excl/.style={box, draw=orange!60!black, fill=orange!6, align=left, text width=3.9cm, font=\footnotesize},
    arr/.style={-{Latex[length=2.5mm]}, thick},
    node distance=9mm and 7mm,
]
\node[src] (s1) {Academic\\survey papers};
\node[src, right=of s1] (s2) {Community-curated\\lists};
\node[src, right=of s2] (s3) {LLM-assisted\\lists};
\node[proc, below=12mm of s2] (cons) {Consolidate equivalent names, then rank by mention count\\[2pt]\footnotesize e.g.\ Gazebo~22, Webots~16, Bullet/PyBullet~16,\\\footnotesize CoppeliaSim/V-REP~13, MuJoCo~12, ODE~10, ROS~2~10, OpenRAVE~9};
\node[proc, below=of cons] (filt) {Filtering criteria (applied in priority order):\\[2pt]open-source availability $\cdot$ repository-based measurability\\$\cdot$ activity \& maintenance (18-month rule) $\cdot$ scope coherence};
\node[final, below=of filt] (fin) {28 selected packages\\[1pt]\footnotesize\mdseries(measured in Chapter~\ref{ch:measurement-results})};
\node[excl, right=10mm of filt] (ex) {\textbf{Excluded:}\\MATLAB/Simulink, Unity, Isaac~Sim, RaiSim (not OSS)\\MRDS, USARSim (inactive)\\GPMP2, CHOMP (scope)};
\draw[arr] (s1) -- (cons);
\draw[arr] (s2) -- (cons);
\draw[arr] (s3) -- (cons);
\draw[arr] (cons) -- (filt);
\draw[arr] (filt) -- (fin);
\draw[arr, dashed, draw=orange!60!black] (filt) -- (ex);
\end{tikzpicture}
\caption{Candidate selection pipeline, from the three discovery sources to the 28 measured packages. The activity criterion is a weighted signal, not a hard cutoff (Section~\ref{subsec:activity-maintenance-status}).}
\label{fig:candidate-selection}
\end{figure}

\begingroup
\singlespacing
\footnotesize
\setlength{\tabcolsep}{4pt}
\begin{longtable}{@{}P{2.5cm}P{3.0cm}llP{2.7cm}@{}}
\caption{Final software set: primary role, repository release date, last commit date, and a representative publication. Repository dates are a May~2026 snapshot; the ``Released'' date is the first release in the measured public repository, which for renamed, re-licensed, or migrated projects (e.g.\ MuJoCo, PhysX, Webots) reflects that later event rather than the project's original creation (Section~\ref{sec:repository-project-metadata}). ``---'' marks packages for which no representative publication was selected; the listed publications are the most representative reference available and are not necessarily peer reviewed.}
\label{tab:final-software-set}\\
\toprule
\textbf{Software} & \textbf{Primary role} & \textbf{Released} & \textbf{Updated} & \textbf{Publication} \\
\midrule
\endfirsthead
\toprule
\textbf{Software} & \textbf{Primary role} & \textbf{Released} & \textbf{Updated} & \textbf{Publication} \\
\midrule
\endhead
\bottomrule
\endlastfoot
\href{https://microsoft.github.io/AirSim}{AirSim} & Simulator (autonomous vehicles) & 2017-02-13 & 2026-03-15 & \citet{sw_airsim} \\
\href{https://pybullet.org}{Bullet/PyBullet} & Physics engine & 2006-05-25 & 2025-10-21 & \citet{sw_pybullet} \\
\href{https://carla.org}{CARLA} & Simulator (autonomous driving) & 2017-03-03 & 2026-05-11 & \citet{sw_carla} \\
\href{https://www.coppeliarobotics.com}{CoppeliaSim (V-REP)} & Simulator (general robotics) & 2019-09-27 & 2026-05-11 & \citet{sw_coppeliasim} \\
\href{https://dartsim.github.io}{DART} & Dynamics library & 2011-04-01 & 2026-05-10 & \citet{sw_dart} \\
\href{https://drake.mit.edu}{Drake} & Toolbox (dynamics, planning, control) & 2010-08-03 & 2026-05-11 & \citet{sw_drake} \\
\href{https://uzh-rpg.github.io/flightmare}{Flightmare} & Simulator (quadrotor) & 2020-07-14 & 2023-05-15 & \citet{sw_flightmare} \\
\href{https://gazebosim.org}{Gazebo} & Simulator (general robotics) & 2014-04-07 & 2026-05-09 & \citet{sw_gazebo} \\
\href{http://graspit-simulator.github.io}{GraspIt!} & Manipulation (grasping) & 2010-10-01 & 2020-07-10 & \citet{sw_graspit} \\
\href{https://aihabitat.org}{Habitat} & Simulator (embodied AI) & 2019-02-25 & 2026-05-07 & \citet{sw_habitat} \\
\href{https://morse-simulator.github.io}{MORSE} & Simulator (academic robotics) & 2009-10-30 & 2022-03-22 & \citet{sw_morse} \\
\href{https://moveit.ai}{MoveIt!} & Manipulation (motion planning) & 2011-10-11 & 2026-03-23 & \citet{sw_moveit} \\
\href{https://www.mrpt.org}{MRPT} & Library (mobile robotics) & 2010-06-03 & 2026-05-10 & --- \\
\href{https://mujoco.org}{MuJoCo} & Physics engine & 2021-10-18 & 2026-05-11 & \citet{sw_mujoco} \\
\href{https://github.com/JulioJerez/newton-dynamics}{Newton Dynamics} & Physics engine & 2013-05-06 & 2026-05-11 & --- \\
\href{https://www.ode.org}{ODE (Open Dynamics Engine)} & Physics engine & 2001-03-04 & 2026-04-14 & --- \\
\href{https://ompl.kavrakilab.org}{OMPL} & Motion planning library & 2010-04-26 & 2026-05-08 & \citet{sw_ompl} \\
\href{http://openrave.org}{OpenRAVE} & Planning framework & 2009-08-15 & 2024-08-16 & \citet{sw_openrave} \\
\href{https://www.openrtm.org}{OpenRTM-aist} & Middleware (RT components) & 2005-05-12 & 2026-05-11 & \citet{sw_openrtm} \\
\href{https://www.orocos.org}{OROCOS} & Control framework & 2007-02-13 & 2026-05-07 & \citet{sw_orocos} \\
\href{https://developer.nvidia.com/physx-sdk}{PhysX (open-source SDK)} & Physics engine & 2022-11-08 & 2026-04-13 & --- \\
\href{https://stack-of-tasks.github.io/pinocchio}{Pinocchio} & Dynamics library & 2014-10-07 & 2026-05-11 & \citet{sw_pinocchio} \\
\href{https://projectchrono.org}{Project Chrono} & Multi-physics simulation & 2010-02-23 & 2026-05-11 & \citet{sw_chrono} \\
\href{https://ros.org}{ROS~2} & Middleware and SDK & 2015-02-24 & 2026-04-29 & \citet{sw_ros2} \\
\href{https://simtk.org/projects/simbody}{Simbody} & Multibody dynamics library & 2005-07-26 & 2026-03-19 & \citet{sw_simbody} \\
\href{https://www.sofa-framework.org}{SOFA} & Simulation framework & 2005-12-08 & 2026-05-08 & \citet{sw_sofa} \\
\href{https://cyberbotics.com}{Webots} & Simulator (general robotics) & 2018-11-05 & 2026-05-04 & \citet{sw_webots} \\
\href{https://www.yarp.it}{YARP} & Middleware (humanoid robotics) & 2006-03-10 & 2026-04-25 & \citet{sw_yarp} \\
\end{longtable}
\endgroup

The role column is necessary because the packages are not direct substitutes; Chapter~\ref{ch:selected-software-commonality} interprets the boundary cases.

\section{Measurement Procedure}
\label{sec:measurement-procedure}

The measurement procedure follows the Domain~X method for collecting structured evidence about software packages in a selected domain~\citep{Smith2021}. For each of the 28 selected packages, the study gathered the public evidence described in Section~\ref{subsec:repository-based-measurability}, and the measurement record identifies the specific repository or repositories measured.

\subsection{Measurement Template}
\label{subsec:measurement-template}

The primary measurement instrument is the measurement template from Appendix~B of \citet{Smith2021}; the version used in this study is reproduced in Appendix~\ref{appendix:measurement-template}. The template contains approximately 108 fields when supplementary entries (free-text ``additional comments'' fields after each section and the GitHub-style and GitStats/SCC repository metric panels) are counted alongside the substantive questions. The substantive questions are grouped into the following sections:

\begin{enumerate}
    \item \textbf{Summary Information} (17 questions): software name, URL, affiliation, purpose, number of developers, funding, initial release date, last commit date, status, license, platforms, software category, development model, publications, source code URL, programming languages, and evidence of performance consideration.
    \item \textbf{Installability} (14 questions): installation instructions (presence, one-place organization, linear ordering, completeness, and assumed pre-installed dependencies), compatible operating system versions, installation automation, error handling, a validation procedure, the number of steps, the operating system used, dependency guidance and version listing, and uninstall behaviour (problems and residue).
    \item \textbf{Correctness and Verifiability} (8 questions): requirements or theory documentation, correctness techniques, tutorial availability and quality, unit tests, and continuous integration evidence.
    \item \textbf{Surface Reliability} (5 questions): installation and tutorial breakage, error messages, and recoverability.
    \item \textbf{Surface Robustness} (2 questions): handling of unexpected input and edge cases.
    \item \textbf{Surface Usability} (4 questions): tutorial, user manual, user characteristics, support model.
    \item \textbf{Maintainability} (7 questions): version number, contribution guidance, available artifacts, issue tracking, issue closure rate, comment percentage, and version control.
    \item \textbf{Reusability} (2 questions): number of code files and API documentation.
    \item \textbf{Surface Understandability} (8 questions): code formatting, coding standards, identifier quality, constant usage, comment clarity, algorithm references, parameter ordering, and modularization.
    \item \textbf{Visibility and Transparency} (4 questions): development process definition, process documentation, development environment documentation, and release notes.
\end{enumerate}

Each quality section concludes with an overall impression score on a scale of 1 to 10, assigned using the scoring guidelines of the impression calculator in Appendix~C of \citet{Smith2021}. The impression scores summarize more than the recorded checklist answers alone: they also reflect observations made during measurement, such as the quality and currency of documentation, the clarity of error messages, and the overall installation and tutorial experience, which the mostly binary checklist fields do not capture. Two packages with identical checklist answers can therefore receive different impression scores; Chapter~\ref{ch:threats-to-validity} discusses the subjectivity this introduces.

\subsection{Repository Metrics}
\label{subsec:repository-metrics}

Repository metrics are collected separately from the quality impression scores. Three categories of repository data are gathered:

\begin{enumerate}
    \item \textbf{GitHub-style metrics:} stars, forks, watchers, open pull requests, closed pull requests, number of developers, open issues, closed issues, initial release date, and last commit date.
    \item \textbf{GitStats-style metrics:} number of text-based files, number of binary files, total lines, lines added, lines deleted, total commits, commits by year (last 5 years), and commits by month (last 12 months).
    \item \textbf{SCC-style metrics:} number of text-based files, total lines, code lines, comment lines, and blank lines.
\end{enumerate}

From these raw data, three processed measures are calculated:

\begin{enumerate}
    \item \textbf{Repository-activity flag:} computed under the 18-month commit-or-release rule (Section~\ref{subsec:activity-maintenance-status}), with \emph{alive} and \emph{dead} recorded as repository-event values rather than complete maintenance judgments.
    \item \textbf{Percentage of identified issues that are closed:} computed as closed issues divided by total identified issues, using GitHub issue data where available.
    \item \textbf{Percentage of code that is comments:} computed with the SCC tool's line-based method, as comment lines divided by total lines.
\end{enumerate}

\subsection{Measurement Execution}
\label{subsec:measurement-execution}

The Domain~X methodology recommends performing installation-based measurements in a controlled environment, typically a virtual machine, to ensure a clean and reproducible starting state~\citep{Smith2021}. For this report, installation measurements were performed on a Linux-based virtual machine. The operating system, installation steps, dependencies, and any issues encountered were recorded for each package.

The methodology also recommends a time budget to keep the overall measurement workload feasible. The full assessment is estimated at roughly 173 person-hours per domain, against a stated goal of about one person-month (160 hours), with 1 to 4 hours allocated per package for template completion and up to 2 hours for installation~\citep{Smith2021}. Measurement was conducted during May 2026. All time-sensitive data (last commit dates, issue counts, stars, forks, etc.) should be understood as snapshots taken during this period.

One further interpretation rule applies: \textbf{evidence that performance was considered} is assessed by searching software artifacts for explicit references to speed, storage, throughput, performance optimization, parallelism, multi-core processing, or similar considerations.

\subsection{Ranking with AHP}
\label{subsec:ahp-ranking}

After the quality measures are collected, the Domain~X methodology ranks the packages with the Analytic Hierarchy Process (AHP), as defined in Section~\ref{sec:ahp-background}, Equations~\eqref{eq:ahp-pairwise}--\eqref{eq:ahp-aggregate}.

In this report, AHP is applied to the nine quality impression scores for the 28 packages. Because no criterion-weight elicitation was carried out, the qualities are weighted equally ($c_k = 1/9$), so the overall priority of Equation~\eqref{eq:ahp-aggregate} reduces to the mean of the nine per-quality priorities. Chapter~\ref{ch:measurement-results} reports the resulting equal-weight AHP ranking alongside the per-package quality-mean ranking and the comparison with community indicators. The robustness of this ranking to the precise impression scores is assessed by perturbing each score by a random amount of up to $\pm 10\%$ and recomputing the ranking, following the sensitivity check of the LBM assessment~\citep{Smith2024LBM}.

\section{Commonality Analysis Procedure}
\label{sec:commonality-analysis-procedure}

The commonality analysis identifies shared and variable capabilities across the selected software packages, treating them as a software family (Section~\ref{sec:commonality-analysis-background}). It supplements the repository-based metrics by describing what the packages do and how their roles differ, providing context for interpreting the measurement results reported in Chapter~\ref{ch:measurement-results}.

Following the Domain~X methodology, the analysis proceeds in six steps:

\begin{enumerate}
    \item \textbf{Write an introduction} to the domain and the software family.
    \item \textbf{Write an overview of domain knowledge,} summarizing the key concepts and terminology relevant to MPS software.
    \item \textbf{List commonalities:} identify goals, assumptions, functions, artifacts, and domain concepts at the level where they hold: family-wide commonalities across all selected packages and sub-family commonalities within role-based groups.
    \item \textbf{List variabilities:} identify dimensions on which family members differ, including role-specific capabilities and implementation choices such as software role, supported platforms, licensing, recorded programming languages, documentation style, and intended use cases.
    \item \textbf{List parameters of variation:} for each variability, describe the possible values.
    \item \textbf{Add terminology, definitions, and acronyms} that are standard in the domain, to ensure that the analysis is accessible to readers outside robotics.
\end{enumerate}

The procedure first groups the selected packages into broad categories based on their primary roles: robotics simulators, physics and dynamics engines, and planning or middleware packages. It then separates family-wide commonalities, sub-family commonalities, and capabilities that vary by role, such as simulation loops, planning workflows, dynamics computation, and ecosystem integration.

Because the selected packages span multiple roles, the commonality analysis does not force all packages into a single comparison template. The analysis is based on checked evidence from documentation, repositories, publications, and other reliable project materials. Capabilities that could not be verified against these materials are not reported.

\section{Interview Component}
\label{sec:interview-component}

The Domain~X methodology includes a developer survey component designed to capture information that is not visible in public repositories: the development process as experienced by practitioners, the pain points they encounter, their perceptions of software quality threats, and the strategies they use to address those threats~\citep{Smith2021}. The methodology prescribes semi-structured interviews based on a list of 18 questions covering developer background, project organization, user understanding, development obstacles (past and present), documentation practices, and specific software qualities including maintainability, understandability, usability, and reproducibility.

The interview component follows a McMaster Research Ethics Board (MREB) protocol. The study received MREB ethics clearance under project~\#8215 on 2026-06-22; recruitment of developers began after clearance. Only evidence supported by approved records and consented summaries is reported.
Interview participants are software developers, maintainers, or programmers associated with the studied packages. The interview guide is reproduced in Appendix~\ref{appendix:interview-materials}.

Chapter~\ref{ch:developer-interviews} addresses the developer interview component and research questions RQ6--RQ9.
% TODO: MREB clearance granted (#8215, 2026-06-22); once interviews are conducted, report the developer findings in Chapter~\ref{ch:developer-interviews} (do not fabricate findings until real interview records exist).

The methodology recommends sending interview requests to developers from all packages in the study set to avoid selection bias. Prior applications of the methodology report response rates roughly in the 15\%--30\% range~\citep{Smith2021,Smith2024MI}; the exact rate for this report will depend on outreach timing and the response practices of the contacted developers. Interview contacts are typically identified through project websites, code repositories, publications, or institutional biography pages.

Chapter~\ref{ch:selected-software-commonality} applies the selection and commonality-analysis procedures to the resulting 28-package study set. It defines the family and its role-based sub-families, then maps each package to the variability values used later to interpret the measurements.
