\chapter{Recommendations for Future Practice}
\label{ch:recommendations}

The recommendations in this chapter come from the public-artifact measurements, the commonality analysis, and the comparison with prior studies of the state of the practice. They are not a ranking of packages or a prescription that every project adopt the same process. RQ8 and RQ9 ask which practices address developer pain points, but those pain points (RQ6) depend on the interview component (Chapter~\ref{ch:developer-interviews}); the recommendations therefore rest on artifact evidence alone and should be revisited once interview evidence is available.

\section{Recommendations by Software Quality}
\label{sec:recommendations-by-quality}

Read together, the measurements in Chapter~\ref{ch:measurement-results} suggest one concrete, low-cost improvement for each measured quality. The list below is organized by the nine qualities; where a recommendation is developed further from the maintainer's point of view, it points to Section~\ref{sec:recommendations-maintainers}.

\begin{itemize}
\item \textbf{Installability.} All 28 packages give installation instructions, but many omit the supported operating-system and dependency versions; recording both, and providing at least one tested installation route, would remove the most common source of installation friction (Section~\ref{sec:recommendations-maintainers}).
\item \textbf{Correctness and verifiability.} Test infrastructure is a strength---unit tests are visible for 27 of the 28 packages---so the useful addition is evidence a user can re-run: reference outputs, benchmark results, or worked examples published alongside the tests let users confirm correctness rather than infer it from the presence of tests.
\item \textbf{Surface reliability and robustness.} The packages generally handle ordinary and malformed input in the surface checks, but that behaviour is rarely documented; a short statement of known limitations and expected failure modes would make the observed reliability something users can depend on.
\item \textbf{Surface usability.} Getting-started tutorials and user manuals are universal, so the gap is audience definition: only one package documents its expected users, and a brief statement of intended audience and prerequisites would help users self-select and give tutorial authors a defined reader (Section~\ref{sec:recommendations-maintainers}).
\item \textbf{Maintainability.} Release notes and contribution or code-review guidance are already present in most repositories (27 and 26 of 28); the few projects without them would most improve maintainability by adopting both, since these are the artifacts that let external contributors participate without direct contact with the original team.
\item \textbf{Reusability.} Every package documents its API, but only 12 of 28 record a documentation-generation tool such as Doxygen or Sphinx; generating API documentation from the source, rather than maintaining it by hand, keeps it current as the interface evolves and lowers the cost of reuse.
\item \textbf{Surface understandability.} All 28 packages format their code consistently, yet only half name an explicit coding standard; naming the standard, or publishing the formatter configuration, turns an implicit convention into a checkable artifact for new contributors (Section~\ref{sec:recommendations-maintainers}).
\item \textbf{Visibility and transparency.} Continuous-integration evidence and a defined development process are visible for most packages, but a short note stating what the automation verifies---builds, tests, documentation, or releases---would make that evidence interpretable rather than merely present.
\end{itemize}

\section{Recommendations for MPS Software Users}
\label{sec:recommendations-users}

Users should choose packages by task fit before quality score. The selected packages serve different roles and are not direct substitutes, so a highly ranked simulator may be the wrong choice for a project that needs a planning library. The commonality analysis in Chapter~\ref{ch:selected-software-commonality} should therefore be read alongside the measurement results in Chapter~\ref{ch:measurement-results}.

Users should also treat public artifacts as evidence of adoption risk. Installation instructions, tutorials, API documentation, issue trackers, and release notes reduce the effort needed to evaluate and maintain a package; sparse documentation, weak release history, or inactive repositories raise it. Popularity indicators deserve caution: Section~\ref{sec:quality-vs-community} shows that GitHub stars and measured engineering quality diverge for several packages, so a star count is not a substitute for checking the artifacts directly.

\section{Recommendations for MPS Software Maintainers}
\label{sec:recommendations-maintainers}

Maintainers should make development practices visible. The measured MPS projects often provide installation guides, tutorials, API references, and issue trackers, but formal and explanatory artifacts are less consistent. Coding standards, user assumptions, design rationale, and requirements-style information are useful because MPS packages are widely adopted by users and contributors who were not part of the original project.

The measurements quantify the three gaps flagged in Section~\ref{sec:recommendations-by-quality}: only one of the 28 packages documents its expected user characteristics, half (14 of 28) do not name a coding standard despite formatting consistently, and compatible operating-system versions are listed by only 61\% of the packages and required dependency versions by 68\%---the main source of the low installability scores in Section~\ref{sec:installability-results}.

Maintainers should also keep repository identity legible. Several measured packages distribute development across repositories or have migrated repositories; Section~\ref{sec:repository-project-metadata} documents PhysX, OROCOS, and Newton Dynamics as examples. A prominent pointer from old locations to the current development home keeps community signals such as stars, forks, and issue history interpretable.

\section[Recommendations for Future State-of-the-Practice Studies]{Recommendations for Future\\ State-of-the-Practice Studies}
\label{sec:recommendations-sop}

Future studies of the state of the practice in heterogeneous software domains should separate domain membership from software role. Recording each package's role through commonality analysis helps avoid treating related packages as interchangeable products.

Future studies should also keep public-artifact evidence separate from developer-reported evidence. Repositories can reveal documentation, issue tracking, releases, testing evidence, and development visibility. They reveal less about internal decision making, informal review, funding constraints, and developer pain points. Those topics require interviews or other qualitative evidence, and claims about them should be tied to the interview records or consented summaries that support them.

Finally, future studies should record at measurement time the observations behind each impression score. Because the scores summarize more than the binary checklist answers (Section~\ref{subsec:measurement-template}), they carry the subjectivity discussed in Chapter~\ref{ch:threats-to-validity}. A short per-package note on which observations drove each score would make the scores easier to audit and reproduce.
