\chapter{Threats to Validity}
\label{ch:threats-to-validity}

Any study based on publicly available artifacts is constrained by what those artifacts can reveal. The threats below are organized by threat type: reliability, construct validity, internal validity, and external validity.

\section{Reliability}
\label{sec:threat-reliability}

All measurements are based on public evidence collected in May 2026 and reflect a snapshot of each repository at that moment (Section~\ref{sec:open-source-repository-measurement}); the same project measured earlier or later might score differently. Some quality questions also require judgement, such as whether documentation is adequate or project metadata complete enough to support a score, and these were performed by a single assessor without an independent inter-rater check. Because one assessor measured all 28 packages in sequence, the criteria may also have been applied slightly differently as familiarity with the template grew---a learning or drift effect that a single rater cannot rule out, and that re-checking early packages against a fixed rubric would reduce but not eliminate. The rationale is recorded in the measurement template (Section~\ref{subsec:measurement-template}), but the impression scores summarize observations beyond the recorded checklist answers, so the recorded data alone cannot fully reconstruct them. The time budget of one to four hours per package bounds the depth of every check; the surface measures are shallow by design.

\section{Construct Validity}
\label{sec:threat-construct-validity}

The measurements use public artifacts as evidence for software qualities, a practical proxy that does not capture every aspect of the underlying constructs: visible testing evidence does not prove that a package is correct, and visible issue-tracking activity does not fully describe maintainability. Mention counts likewise measure how often a package appears across candidate sources, not how good the software is (Section~\ref{sec:mention-count-ranking}). The choice of quality categories is itself a construct threat: the nine categories applied here (Section~\ref{subsec:measurement-template}) are one reasonable decomposition of ``software quality'' rather than the only one, and because the impression scores involve assessor judgement (Section~\ref{sec:threat-reliability}), a reader who defines or weights quality differently could reach different conclusions.

\section{Internal Validity}
\label{sec:threat-internal-validity}

The candidate set was built from academic survey papers, community-curated lists, and LLM-assisted discovery, with the LLM lists supplementing rather than driving the other two sources. Those lists reflect the training data of the model rather than an independent survey of the MPS domain, so a package unknown to the model could be absent from its output. The filtering decisions described in Section~\ref{sec:filtering-criteria} further shape the measured set, since a package can be excluded for being out of scope, closed source, inactive, or not measurable from public repositories.

The overall ranking also depends on how the per-quality scores are combined. The report uses an equal-weight aggregation (Chapter~\ref{ch:measurement-results}), which treats all nine qualities as equally important; a different, application-specific weighting could reorder the middle of the ranking. The sensitivity analysis reported in Chapter~\ref{ch:measurement-results} perturbs the scores and re-computes the ranking, and finds the top and bottom stable, but the equal-weight choice itself remains a modelling decision rather than an empirical finding.

\section{External Validity}
\label{sec:threat-external-validity}

The selected set covers open-source MPS software measurable through public artifacts. Proprietary platforms, discontinued projects, and narrowly scoped libraries may differ from the measured packages, so the results describe the selected open-source MPS sample, not all robotics software. Because the packages play different roles, comparisons are evidence about development practice across a domain rather than direct product-to-product comparisons.

The installation-based measurements were performed on a Linux virtual machine, so the installability, surface reliability, and surface robustness results reflect Linux behaviour only. The same packages may install and behave differently on Windows or macOS, and installation routes available at measurement time, such as distribution packages, may be absent on other operating systems or later distribution releases.

\section{Threats to the Planned Interview Component}
\label{sec:threat-interviews}

The developer-interview component (Chapter~\ref{ch:developer-interviews}) is designed but not yet conducted, so the threats below will apply once it is carried out and should be read as limitations of that planned evidence rather than of the artifact measurements above. Participation is voluntary, so the developers who respond may not represent the domain: those who agree to be interviewed, or who maintain more visible projects, can differ systematically from those who do not, giving rise to self-selection and non-response bias. The number of interviews feasible within the project timeline is small, so the findings will illustrate rather than establish domain-wide patterns. Because the interviews are semi-structured and conducted by the researcher, the wording and follow-up questions can steer answers (interviewer bias), and developers may describe their practices more favourably than they are (self-report and social-desirability bias). These threats will be reduced, but not removed, by following the approved interview guide, separating reported current practice from aspiration, and reporting the number of participants and the packages they represent.
